---
title: "Chikv | Statistics and Plots before filtering | Merged"
author: 'Giorgio Gonnella'
date: "`r Sys.Date()`"
output:
  html_document: default
params:
  prjpath: "/srv/immunology/chikv-s5/"
  analysis: "1"
  step: "02.V.filtered"
  input_step: "02.P.filter"
  input_sfx: "_so_filtered.rds"
  samples_sheet: "metadata/samples.tsv"
  sample_id_col: "SampleID"
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(message=F, warning=F, echo=F)
withr::with_dir(here::here(), { source("renv/activate.R") })

library(Seurat)
library(tictoc)
library(dplyr)
library(ggbeeswarm)
library(ggExtra)
library(matrixStats)
library(rmdreportdeck)
library(baia.seurat.helpers)
reportdeck_setup()
knitr::opts_chunk$set(cache=F)

analysis_root <- here::here(paste0("analysis-", params$analysis,"/"))
samples_sheet <- here::here(params$samples_sheet)

input_step_dir <- paste0(analysis_root, "steps/", params$input_step, "/")
input_var_dir <- paste0(input_step_dir, "results/vars/")
if (!dir.exists(input_var_dir)) {
  stop("Input variable directory does not exist: ", input_var_dir)
}

this_step_dir <- paste0(analysis_root, "steps/", params$step, "/")
if (!dir.exists(this_step_dir)) {
  stop("Current step directory does not exist: ", this_step_dir)
}

```

# Goal

Visualize quality control metrics across all samples after filtering, merged into a single object.
Confirms that the applied filtering thresholds effectively removed low-quality cells and provides
per-sample and merged QC plots for the cleaned dataset.

# Load Seurat objects

```{r read input data}
samples_metadata <- read_samples_sheet(samples_sheet,
                                       column=params$sample_id_col)
sampleIDs <- baia.seurat.helpers::get_sample_IDs(
  samples_metadata,
  column=params$sample_id_col
)

so_list = list()
for (sampleID in sampleIDs) {
  filename = paste0(input_var_dir, sampleID, params$input_sfx)
  print(filename)
  so_list[[sampleID]] <- readRDS(filename)
}
```

# Filtering information by sample

```{r, asis=TRUE}
for (sampleID in sampleIDs) {
  so <- so_list[[sampleID]]
  print("")
  print(paste0("Sample: ", sampleID))
  print(paste0("nCells before filtering: ", so@misc$qc_filtering$nCells))
  print(paste0("nCells after filtering: ", ncol(so)))
  print(paste0("nGenes filter: ",
               so@misc$qc_filtering$nGenes))
  print(paste0("nCells failing nGenes filter: ",
               so@misc$qc_filtering$nCells_failing_nGenes))
  print(paste0("nUMI filter: ", so@misc$qc_filtering$nUMI))
  print(paste0("nCells failing nUMIs filter: ",
                so@misc$qc_filtering$nCells_failing_nUMIs))
  print(paste0("nMito filter: ", so@misc$qc_filtering$pmito))
  print(paste0("nCells failing pmito filter: ",
                so@misc$qc_filtering$nCells_failing_pmito))
  print(paste0("nCells failing multi-outliers filter: ",
                so@misc$qc_filtering$nCells_failing_multi_outliers))
  print(paste0("multi-outliers filter applied: ",
                so@misc$qc_filtering$multi_outliers_applied))
  print(paste0("nCells failing celltype filter: ",
                so@misc$qc_filtering$nCells_failing_SingleR))
  print("")
}
```

```{r, asis=TRUE}
for (sampleID in sampleIDs) {
  so <- so_list[[sampleID]]
  print("")
  print("celltypes main after celltype filter:")
  print(table(so$pruned.labels_singleR_main))
  print("celltypes fine after celltype filter:")
  print(table(so$pruned.labels_singleR_fine))
  print("")
  print("")
}
```

# Merge objects into a single Seurat object

```{r merge}
so_merged <- merge(x = so_list[[1]], y = so_list[-1], project = "chikv")
so_merged$sample <- gsub("_so", "", so_merged$orig.ident)
so_merged$mitoRatio = so_merged$percent.mito / 100
so_merged$log10GenesPerUMI <- log10(so_merged$nFeature_RNA) / log10(so_merged$nCount_RNA)
table(so_merged$sample)
summary(so_merged@meta.data)
```

# QC Plots after filtering

## Plots for all samples

```{r plots}
histogram_n_cells(so_merged)
so_merged@meta.data %>% group_by(sample) %>% summarise(n_cells = n())
print(density_plot_n_umis(so_merged, 7))
print(boxplot_n_umis(so_merged))
print(density_plot_n_genes(so_merged, 7))
print(boxplot_n_genes(so_merged))
print(density_plot_mito_ratio(so_merged, 7))
print(dotplot_n_umis_genes_mito(so_merged, 7))
print(density_plot_complexity(so_merged, 7))
```
## Plots divided by sample

```{r highest expressed genes}
for (sampleID in sampleIDs) {
  plot_highest_expressed(so_list[[sampleID]])
}
```

```{r}
for (sampleID in sampleIDs) {
  so_list[[sampleID]]@misc$nGene_th <-
    c(attr(so_list[[sampleID]]@meta.data$qc.nexprs.low, "thresholds")["lower"],
      attr(so_list[[sampleID]]@meta.data$qc.nexprs.high, "thresholds")["higher"])
  so_list[[sampleID]]@misc$nUMI_th <-
    c(attr(so_list[[sampleID]]@meta.data$qc.lib.low, "thresholds")["lower"],
      attr(so_list[[sampleID]]@meta.data$qc.lib.high, "thresholds")["higher"])
  so_list[[sampleID]]@misc$nMito_th <-
    attr(so_list[[sampleID]]@meta.data$qc.mito.high, "thresholds")["higher"]
}
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=5}
for (sampleID in sampleIDs) {
  so <- so_list[[sampleID]]
  qcstats <- so@meta.data
  nGene_th <- so_list[[sampleID]]@misc$nGene_th
  nUMI_th <- so_list[[sampleID]]@misc$nUMI_th
  nMito_th <- so_list[[sampleID]]@misc$nMito_th

  pmito <- ggplot(qcstats, aes(x = "", y = percent.mito*100))+
     geom_quasirandom(width = .4, aes(color = qc.mito.high), alpha = .7)+
     geom_violin(draw_quantiles = .5, fill = NA, color = "black", width = .4)+
     theme_minimal()+
     scale_color_manual(values = c("steelblue2", "tan2"))+
     guides(color=FALSE)+
     geom_hline(yintercept = nMito_th[["higher"]], linetype ="dashed")+
     xlab('') + ggtitle(paste0("Sample: ", sampleID))
  
  pumi <- ggplot(qcstats, aes(x = "", y = scater_qcstats_sum))+
     geom_quasirandom(width = .4,aes(color = qc.lib.low | qc.lib.high), alpha = .7)+
     geom_violin(draw_quantiles = .5, fill = NA, color = "black", width = .4)+
     theme_minimal()+
     scale_color_manual(values = c("steelblue2", "tan2"))+
     guides(color=FALSE)+
     geom_hline(yintercept = c(nUMI_th[["lower"]], nUMI_th[["higher"]]),
                linetype ="dashed")+
     xlab('') + ggtitle(paste0("Sample: ", sampleID))
  
  pgene <- ggplot(qcstats, aes(x = "", y = scater_qcstats_detected))+
     geom_quasirandom(width = .4, aes(color = qc.nexprs.low | qc.nexprs.high), alpha = .7)+
     geom_violin(draw_quantiles = .5, fill = NA, color = "black", width = .4)+
     theme_minimal()+
     scale_color_manual(values = c("steelblue2", "tan2"))+
     guides(color=FALSE)+
     geom_hline(yintercept = c(nGene_th[["lower"]], nGene_th[["higher"]]),
                linetype ="dashed")+
     xlab('') + ggtitle(paste0("Sample: ", sampleID))
  
  grid.arrange(pmito, pumi, pgene, ncol = 3)
}
```

```{r, fig.align="center"}
for (sampleID in sampleIDs) {
  so <- so_list[[sampleID]]
  qcstats <- so@meta.data
  plt <- ggplot(qcstats,aes(x = scater_qcstats_sum, y = percent.mito*100, color = qc.mito.high))+
     geom_point(alpha = .7)+
     scale_color_manual(values = c("steelblue2", "tan2"))+
     guides(color=FALSE)+
     theme_minimal()+
     geom_vline(xintercept = c(500, 1000), linetype = "dashed", color = "firebrick2")+
     geom_hline(yintercept = 10, linetype = "dashed") + ggtitle(paste0("Sample: ", sampleID))
  print(plt)
}
```
```{r}
for (sampleID in sampleIDs) {
  so <- so_list[[sampleID]]
  p_nUMI_mito <- ggplot(so@meta.data, aes(x = nCount_RNA, y = percent.mito, fill = percent.mito.cl))+
    geom_point(alpha = .6, color = "grey25", shape = 21) +
    theme_minimal() +
    scale_fill_viridis_d() +
    geom_vline(xintercept = c(1000, 5000), linetype = "dashed") +
    scale_x_continuous(breaks = seq(from = 0, to = 1e5, by = 1e4)) +
    geom_hline(yintercept = .1, linetype = "dashed") +
    ggtitle(paste0("Sample: ", sampleID))
  print(ggMarginal(p_nUMI_mito, type = "density"))
}
```

```{r}
for (sampleID in sampleIDs) {
  so <- so_list[[sampleID]]
  qcstats <- so@meta.data
  so@meta.data$umiSum.mito <- qcstats$subsets_Mito_sum
  p <- ggplot(so@meta.data, aes(x = nCount_RNA, y = umiSum.mito, fill = percent.mito.cl)) +
    geom_point(alpha = .6, color = "grey25", shape = 21) +
    theme_minimal() +
    scale_fill_viridis_d() +
    ggtitle(paste0("Sample: ", sampleID))
  grid.arrange(p, p+facet_wrap(~percent.mito.cl, ncol = 2), ncol=2)
}
```

# Session info

```{r session_info, echo=F}
sessionInfo()
```
