---
title: "`r paste('Chikv | Differential expression', c(A='Acute', M6='6-months', NC='Non-chronic', C='Chronic')[params$pair], c(A='Non-chronic vs Chronic', M6='Non-chronic vs Chronic', NC='Acute vs 6-months', C='Acute vs 6-months')[params$pair], sep=' | ')`"
author: "Giorgio Gonnella"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: sandstone
params:
  analysis: "1"
  input_step: "05.P.compute_de"
  input_rds: "so_w_deseq.rds"
  step: "05.V.de_results"
  pair: "A"   # one of: A, M6, NC, C
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=F, warning=F, echo=F)
withr::with_dir(here::here(), { source("renv/activate.R") })
withr::with_dir(here::here(), { source("helpers/de_helpers.R") })
withr::with_dir(here::here(), { source("helpers/de_viz_helpers.R") })

library(Seurat)
library(dplyr)
require(data.table)
library(EnhancedVolcano)
library(rmdreportdeck)
reportdeck_setup()
knitr::opts_chunk$set(cache=F)

analysis_root <- here::here(paste0("analysis-", params$analysis,"/"))

input_step_dir <- paste0(analysis_root, "steps/", params$input_step, "/")
input_var_dir <- paste0(input_step_dir, "results/vars/")
if (!dir.exists(input_var_dir))
  stop("Input variable directory does not exist: ", input_var_dir)

this_step_dir <- paste0(analysis_root, "steps/", params$step, "/")
if (!dir.exists(this_step_dir))
  stop("Current step directory does not exist: ", this_step_dir)

lbl_map    <- list(A = "C_vs_NC", M6 = "C_vs_NC", NC = "A_vs_6M", C = "A_vs_6M")
sister_map <- list(A = "M6",     M6 = "A",       NC = "C",       C = "NC")
if (!params$pair %in% names(lbl_map))
  stop("Unknown pair: ", params$pair,
       ". Must be one of: ", paste(names(lbl_map), collapse=", "))

n      <- params$pair
lbl    <- lbl_map[[n]]
pfx    <- paste0(n, ".", lbl)   # e.g. "A.C_vs_NC"
sister <- sister_map[[n]]
contrast_groups <- strsplit(lbl, "_vs_")[[1]]
findmarkers_ident1 <- list(A = "Chronic", M6 = "Chronic", NC = "A", C = "A")[[n]]
findmarkers_ident2 <- list(A = "Non-Chronic", M6 = "Non-Chronic", NC = "M6", C = "M6")[[n]]
findmarkers_call <- paste0('FindMarkers(..., ident.1 = "', findmarkers_ident1,
                           '", ident.2 = "', findmarkers_ident2, '")')
log2fc_positive_group <- findmarkers_ident1
log2fc_negative_group <- findmarkers_ident2

p_adj_filt_tsv_dir <- paste0(this_step_dir, "results/tables/p_adj_filt/")
dir.create(p_adj_filt_tsv_dir, showWarnings = FALSE, recursive = TRUE)
vulcano_outdir <- paste0(this_step_dir, "results/plots/vulcano/")
dir.create(vulcano_outdir, recursive = TRUE, showWarnings = FALSE)

strsanitize <- function(s) gsub(" ", "_", gsub("[^[:alnum:]]", "_", s))

h <- function(level, text) writeLines(c("", paste0(strrep("#", level), " ", text), ""))

deg_table_cols <- c("gene", "log2fc_group_1", "avg_expr_group_1",
                    "log2fc_group_2", "avg_expr_group_2", "higher_in",
                    "avg_log2FC", "pct.1", "pct.2", "p_val", "p_val_adj")
deg_tsv_cols <- c("cmp", "ct", deg_table_cols)
available_cols <- function(x, cols) cols[cols %in% colnames(x)]
add_de_report_expression_cols <- function(x) {
  if ("avg_log2expr_group_1" %in% colnames(x)) {
    x$avg_expr_group_1 <- 2 ^ x$avg_log2expr_group_1
  }
  if ("avg_log2expr_group_2" %in% colnames(x)) {
    x$avg_expr_group_2 <- 2 ^ x$avg_log2expr_group_2
  }
  x
}
```

```{r load_input_data}
so <- readRDS(file = paste0(input_var_dir, params$input_rds))
```

# Goal

<style>
table caption {
  color: black;
  font-weight: bold;
}
</style>

Visualize DESeq2 differential expression results for a given pairwise comparison as volcano plots
and tables, one cell type at a time. Highlights significantly differentially expressed genes
(adjusted p < 0.05, |log2FC| > 0.58) and collects cell types with no significant DEGs at the end.

For this report, the DE call is `r findmarkers_call`, which means that positive
`avg_log2FC` values are higher in `r findmarkers_ident1` and negative `avg_log2FC`
values are higher in `r findmarkers_ident2`. For cell-type-specific tests, the
same direction applies to the corresponding cell-type identities (for example,
`r paste0(findmarkers_ident1, "_<cell type>")` versus
`r paste0(findmarkers_ident2, "_<cell type>")`).

**Tables:** The asset-bar TSV download for each cell type contains all significant DEGs. The
inline tables are shortened to the top 20 rows, sorted by adjusted p-value from smallest to
largest. The average expression columns are shown on the expression scale, while
`avg_log2FC` remains on the standard log2 fold-change scale.

**Sister-group significant genes:** Each cell type section also provides a downloadable TSV of
genes that are significant in the sister group (`r sister`) for the same comparison axis but may
not reach significance here. This allows cross-referencing genes with consistent directional
signal across both groups.

```{r plots_and_tables, results='asis'}
de          <- so@misc$de[[n]]
de_sister   <- so@misc$de[[sister]]
celltypes   <- names(de$pval_filt)
total_cts   <- celltypes[celltypes == "Total"]
celltypes   <- setdiff(celltypes, total_cts)
no_deg_cts  <- character(0)
fig_counter <- 0L
table_n     <- 20L

h(1, "Significant DEG counts")
deg_count_plot <- de_direction_count_plot(
  de$pval_filt,
  positive_condition = log2fc_positive_group,
  negative_condition = log2fc_negative_group,
  celltypes = celltypes,
  title = paste0(pfx, " significant DEGs")
)
deg_count_bundle <- report_plot_bundle(
  deg_count_plot,
  block_prefix = paste0("DEGCounts.", pfx),
  suffix = "direction_counts",
  width = 8,
  height = 5,
  plot_dir = vulcano_outdir
)
cat(as.character(report_bundle_asset_bar(
  report_figure_name(paste0("DEGCounts.", pfx), suffix = "direction_counts"),
  deg_count_bundle,
  png_label = "Download PNG",
  pdf_label = "Download PDF"
)), "\n")
cat(
  paste0('<img src="', deg_count_bundle$png_href, '" width="768" />'),
  "\n\n"
)

get_sister_sig_table <- function(ct) {
  cts <- strsanitize(ct)
  sister_sig_ct <- de_sister$pval_filt[[ct]]
  all_ct <- de$all[[ct]]

  if (is.null(sister_sig_ct) || nrow(sister_sig_ct) == 0 ||
      is.null(all_ct) || nrow(all_ct) == 0) {
    return(NULL)
  }

  all_ct$gene <- rownames(all_ct)
  all_ct <- add_de_report_expression_cols(all_ct)
  sister_gene_order <- rownames(sister_sig_ct)[order(sister_sig_ct$p_val_adj)]
  cross_data <- all_ct[match(sister_gene_order, all_ct$gene), , drop = FALSE]
  cross_data <- cross_data[!is.na(cross_data$gene), , drop = FALSE]

  if (nrow(cross_data) == 0) {
    return(NULL)
  }

  cross_data <- cross_data[, available_cols(cross_data, deg_table_cols)]
  table_caption <- paste0(pfx, " \u2014 ", ct, " \u2014 Significant in ", sister)
  if (nrow(cross_data) > table_n) {
    table_caption <- paste0(
      table_caption,
      " \u2014 Top ", table_n, " by sister-pair adjusted p-value"
    )
  }

  list(
    data = cross_data,
    tsv_fname = paste0(pfx, ".", cts, ".sig_in_", sister, ".tsv"),
    caption = table_caption,
    suffix = paste0(cts, ".sig_in_", sister)
  )
}

render_sister_sig_table <- function(ct, table_block_prefix, emit_total_subheader = FALSE) {
  sister_table <- get_sister_sig_table(ct)
  if (is.null(sister_table)) {
    return(FALSE)
  }

  if (emit_total_subheader) {
    h(2, paste0("Significant in ", sister))
  }

  cat(as.character(report_asset_bar(
    report_table_name(table_block_prefix, sister_table$suffix),
    report_download_link(paste0("Download TSV (sig. in ", sister, ")"),
                         sister_table$tsv_fname,
                         report_text_data_uri(report_tsv_text(sister_table$data)))
  )), "\n")

  print(knitr::kable(
    format_de_display_table(head(sister_table$data, table_n)),
    digits = 4,
    caption = sister_table$caption,
    row.names = FALSE,
    escape = FALSE
  ))

  TRUE
}

render_deg_section <- function(ct, heading_level = 2L, heading_text = ct,
                               emit_heading = TRUE, total_tables = FALSE) {
  data_ct  <- de$pval_filt[[ct]]
  cts      <- strsanitize(ct)
  has_data <- !is.null(data_ct) && nrow(data_ct) > 0

  # TSV
  if (has_data) {
    data_ct$cmp  <- lbl
    data_ct$ct   <- ct
    data_ct$gene <- rownames(data_ct)
    data_ct <- add_de_report_expression_cols(data_ct)
    fwrite(data_ct[, available_cols(data_ct, deg_tsv_cols)],
           file = paste0(p_adj_filt_tsv_dir, pfx, ".", cts, ".tsv"),
           sep = "\t", quote = FALSE)
  }

  if (!has_data) {
    return(FALSE)
  }

  fig_counter <<- fig_counter + 1L

  if (emit_heading) {
    h(heading_level, heading_text)
  }

  png_path <- paste0(vulcano_outdir, pfx, ".", cts, ".png")
  plt <- EnhancedVolcano(data_ct, lab = rownames(data_ct),
                         x = 'avg_log2FC', y = 'p_val_adj',
                         FCcutoff = 0.58, pCutoff = 0.05) +
    ggtitle(paste0(n, " | ", lbl, " | ", ct))
  figure_id <- paste0(fig_counter, ".", pfx)
  plot_bundle <- report_plot_bundle(
    plt,
    block_prefix = figure_id,
    suffix = cts,
    width = 7,
    height = 5,
    plot_dir = vulcano_outdir
  )
  ggsave(png_path, plt, width=7, height=5)
  cat(as.character(report_bundle_asset_bar(
    report_figure_name(figure_id, suffix = cts),
    plot_bundle,
    png_label = "Download PNG",
    pdf_label = "Download PDF"
  )), "\n")
  cat(
    paste0(
      '<img src="', plot_bundle$png_href, '" width="672" />'
    ),
    "\n\n"
  )

  table_data <- data_ct[order(data_ct$p_val_adj),
                        available_cols(data_ct, deg_table_cols)]
  table_caption <- paste0(pfx, " \u2014 ", ct)
  if (nrow(table_data) > table_n) {
    table_caption <- paste0(table_caption, " \u2014 Top ", table_n, " by adjusted p-value")
  }

  tsv_fname <- paste0(pfx, ".", cts, ".tsv")
  if (total_tables) {
    h(2, "Significant DEGs")
  }
  cat(as.character(report_asset_bar(
    report_table_name(figure_id, cts),
    report_download_link("Download TSV", tsv_fname,
                         report_text_data_uri(report_tsv_text(table_data)))
  )), "\n")

  print(knitr::kable(
    format_de_display_table(head(table_data, table_n)),
    digits = 4,
    caption = table_caption,
    row.names = FALSE,
    escape = FALSE
  ))

  render_sister_sig_table(ct, figure_id, emit_total_subheader = total_tables)
  TRUE
}

h(1, "Total cells")
if (length(total_cts) == 0) {
  writeLines("Total cells were not present in the DE results.")
} else {
  for (ct in total_cts) {
    if (!render_deg_section(ct, emit_heading = FALSE, total_tables = TRUE)) {
      writeLines("No significant DEGs.")
    }
  }
}

h(1, "Cell types with significant DEGs")
for (ct in celltypes) {
  if (!render_deg_section(ct)) {
    no_deg_cts <- c(no_deg_cts, ct)
  }
}

h(1, paste0("Cell types with no significant DEGs in ", n,
            " but significant DEGs in ", sister))
no_deg_also_sister <- character(0)
if (length(no_deg_cts) == 0) {
  writeLines("None")
} else {
  no_sister_tables <- 0L
  for (ct in no_deg_cts) {
    if (!is.null(get_sister_sig_table(ct))) {
      no_sister_tables <- no_sister_tables + 1L
      h(2, ct)
      render_sister_sig_table(ct, paste0("NoSig.", no_sister_tables, ".", pfx))
    } else {
      no_deg_also_sister <- c(no_deg_also_sister, ct)
    }
  }
  if (no_sister_tables == 0L) {
    writeLines("None")
  }
}

h(1, paste0("Cell types with no significant DEGs in ", n, " and ", sister))
if (length(no_deg_also_sister) == 0) {
  writeLines("None")
} else {
  writeLines(paste0("- ", no_deg_also_sister))
}

```

# Session info

```{r}
sessionInfo()
```
