---
title: "`r paste('Chikv | Differential expression', c(A = 'Acute', M6 = '6-months', NC = 'Non-chronic', C = 'Chronic')[params$pair], c(A = 'Non-chronic vs Chronic', M6 = 'Non-chronic vs Chronic', NC = 'Acute vs 6-months', C = 'Acute vs 6-months')[params$pair], sep = ' | ')`"
author: "Giorgio Gonnella"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: sandstone
params:
  analysis: "1"
  input_step: "05.P.compute_de"
  input_rds: "so_w_deseq.rds"
  step: "05.V.de_results"
  pair: "A"
  table_n: 20
  fc_cutoff: 0.58
  p_cutoff: 0.05
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
withr::with_dir(here::here(), { source("renv/activate.R") })

library(Seurat)
library(baia.pseudobulk.deseq)
library(rmdreportdeck)

rmdreportdeck::reportdeck_setup()
knitr::opts_chunk$set(cache = FALSE)

pair_configs <- list(
  A = list(
    lbl = "C_vs_NC",
    sister = "M6",
    positive_condition = "Chronic",
    negative_condition = "Non-Chronic",
    title = "Acute | Chronic vs Non-Chronic"
  ),
  M6 = list(
    lbl = "C_vs_NC",
    sister = "A",
    positive_condition = "Chronic",
    negative_condition = "Non-Chronic",
    title = "6-months | Chronic vs Non-Chronic"
  ),
  NC = list(
    lbl = "A_vs_6M",
    sister = "C",
    positive_condition = "A",
    negative_condition = "M6",
    title = "Non-Chronic | Acute vs 6-months"
  ),
  C = list(
    lbl = "A_vs_6M",
    sister = "NC",
    positive_condition = "A",
    negative_condition = "M6",
    title = "Chronic | Acute vs 6-months"
  )
)

if (!params$pair %in% names(pair_configs)) {
  stop("Unknown pair: ", params$pair,
       ". Must be one of: ", paste(names(pair_configs), collapse = ", "),
       call. = FALSE)
}

cfg <- pair_configs[[params$pair]]
n <- params$pair
lbl <- cfg$lbl
pfx <- paste0(n, ".", lbl)
sister <- cfg$sister
findmarkers_ident1 <- cfg$positive_condition
findmarkers_ident2 <- cfg$negative_condition
findmarkers_call <- paste0('FindMarkers(..., ident.1 = "', findmarkers_ident1,
                           '", ident.2 = "', findmarkers_ident2, '")')

analysis_root <- here::here(paste0("analysis-", params$analysis, "/"))

input_step_dir <- paste0(analysis_root, "steps/", params$input_step, "/")
input_var_dir <- paste0(input_step_dir, "results/vars/")
if (!dir.exists(input_var_dir)) {
  stop("Input variable directory does not exist: ", input_var_dir)
}

this_step_dir <- paste0(analysis_root, "steps/", params$step, "/")
dir.create(this_step_dir, recursive = TRUE, showWarnings = FALSE)

tables_outdir <- paste0(this_step_dir, "results/tables/p_adj_filt/")
plots_outdir <- paste0(this_step_dir, "results/plots/vulcano/")
dir.create(tables_outdir, showWarnings = FALSE, recursive = TRUE)
dir.create(plots_outdir, recursive = TRUE, showWarnings = FALSE)
```

```{r load_input_data}
so <- readRDS(file = paste0(input_var_dir, params$input_rds))
de <- so@misc$de[[n]]
de_sister <- so@misc$de[[sister]]
celltypes <- names(de$pval_filt)
total_cts <- celltypes[celltypes == "Total"]
plot_celltypes <- setdiff(celltypes, total_cts)
```

# Goal

<style>
table caption {
  color: black;
  font-weight: bold;
}
</style>

Visualize DESeq2 differential expression results for a given pairwise comparison as volcano plots
and tables, one cell type at a time. Highlights significantly differentially expressed genes
(adjusted p < `r params$p_cutoff`, |log2FC| > `r params$fc_cutoff`) and collects cell types with no
significant DEGs at the end.

For this report, the DE call is `r findmarkers_call`, which means that positive
`avg_log2FC` values are higher in `r findmarkers_ident1` and negative `avg_log2FC`
values are higher in `r findmarkers_ident2`. For cell-type-specific tests, the
same direction applies to the corresponding cell-type identities (for example,
`r paste0(findmarkers_ident1, "_<cell type>")` versus
`r paste0(findmarkers_ident2, "_<cell type>")`).

**Tables:** The asset-bar TSV download for each cell type contains all significant DEGs. The
inline tables are shortened to the top `r params$table_n` rows, sorted by adjusted p-value from
smallest to largest. The average expression columns are shown on the expression scale, while
`avg_log2FC` remains on the standard log2 fold-change scale.

**Sister-group significant genes:** Each cell type section also provides a downloadable TSV of
genes that are significant in the sister group (`r sister`) for the same comparison axis but may
not reach significance here. This allows cross-referencing genes with consistent directional
signal across both groups.

```{r direction_counts, results='asis'}
de_report_direction_counts_section(
  de = de,
  positive_condition = cfg$positive_condition,
  negative_condition = cfg$negative_condition,
  celltypes = plot_celltypes,
  pfx = pfx,
  plots_outdir = plots_outdir
)
```

```{r plots_and_tables, results='asis'}
de_report_celltype_sections(
  de = de,
  de_sister = de_sister,
  sister = sister,
  pfx = pfx,
  lbl = lbl,
  celltypes = celltypes,
  table_n = params$table_n,
  plots_outdir = plots_outdir,
  tables_outdir = tables_outdir,
  fc_cutoff = params$fc_cutoff,
  p_cutoff = params$p_cutoff
)
```

# Session info

```{r session_info}
sessionInfo()
```
