---
title: "Chikv | Create Seurat 5 object for a sample"
author: "Giorgio Gonnella"
date: "2025-06-25"
output:
  html_document: default
params:
  prjpath: "/srv/immunology/chikv-s5/"
  analysis: "1"
  step: "01.P.create_so"
  sample: NA
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(message=F, warning=F, echo=F)
withr::with_dir(here::here(), { source("renv/activate.R") })

library(SingleCellExperiment)
library(scater)
library(Seurat)
library(robustbase)

if (is.na(params$sample)) {
  stop("Please provide a sample name")
}

analysis_root <- here::here(paste0("analysis-", params$analysis,"/"))
data_dir <- here::here("cellranger/results/")

sample_data_dir <- paste0(data_dir, params$sample, "/")
if (!dir.exists(sample_data_dir)) {
  stop("Sample data directory does not exist: ", sample_data_dir)
}

this_step_dir <- paste0(analysis_root, "steps/", params$step, "/")
if (!dir.exists(this_step_dir)) {
  stop("Current step directory does not exist: ", this_step_dir)
}

var_dir <- paste0(this_step_dir, "results/vars/")
dir.create(var_dir, showWarnings = FALSE, recursive = TRUE)

helpers_dir <- here::here("helpers/")
source(paste0(helpers_dir, "qc_helpers.R"))
```

# Create SO

Create Seurat object from Cellranger raw h5 matrix, in order to identify the
cells where CHIKV genes are present, then use the filtered matrix to identify
the other cells and genes to keep, and filter the Seurat object using them, but
always keeping all infected cells and the CHIKV genes

```{r function_to_extract_counts_of_genes_of_interest}
GENES_OF_INTEREST = c("CHIKVgp1", "CHIKVgp2")

get_goi_counts <- function(barcodes_of_interest, genes_of_interest, so) {
  valid_barcodes <- barcodes_of_interest %in% colnames(so)
  valid_genes <- genes_of_interest %in% rownames(so)
  if (any(valid_barcodes) & any(valid_genes)) {
      data_subset <- so[genes_of_interest[valid_genes],
                        barcodes_of_interest[valid_barcodes]]
      counts <- as.matrix(GetAssayData(data_subset))
  } else {
      counts <- "No matching barcodes or genes found"
  }
  counts
}
```

```{r create_so_with_raw_data}
raw_h5fn = paste0(sample_data_dir, "raw_feature_bc_matrix.h5")
if (!file.exists(raw_h5fn)) {
  stop("No raw h5 file for sample", params$sample,
       " found in directory ", sample_data_dir)
}
raw_counts <- Read10X_h5(filename = raw_h5fn)
so_name <- paste0(params$sample, "_so")
raw_counts_so <- CreateSeuratObject(counts = raw_counts, project = so_name,
                         min.cells = 0, min.features = 0)
rm(raw_counts)
```

```{r identify_cells_with_CHIKV_genes}
raw_counts_goi_so <- subset(raw_counts_so, features = GENES_OF_INTEREST)
goi_sums <- Matrix::colSums(GetAssayData(raw_counts_goi_so))
rm(raw_counts_goi_so)

goi_cells <- names(goi_sums[goi_sums > 0])
rm(goi_sums)
```


```{r create_so_with_filtered_data}
filtered_h5fn = paste0(sample_data_dir, "filtered_feature_bc_matrix.h5")
if (!file.exists(filtered_h5fn)) {
  stop("No filtered h5 file for sample", params$sample,
       " found in directory ", sample_data_dir)
}
filtered_counts <- Read10X_h5(filename = filtered_h5fn)
so_name <- paste0(params$sample, "_so")
so <- CreateSeuratObject(counts = filtered_counts, project = so_name,
                         min.cells = 0, min.features = 0)
rm(filtered_counts)
```

```{r}
if (length(goi_cells) > 0) {
  print("Cells with CHIKV genes in the filtered data:")
  print(get_goi_counts(goi_cells, GENES_OF_INTEREST, so))
  print("Cells with CHIKV genes in the raw data:")
  print(get_goi_counts(goi_cells, GENES_OF_INTEREST, raw_counts_so))
  filtered_cells <- colnames(so)
  filtered_genes <- rownames(so)
  all_cells <- union(filtered_cells, goi_cells)
  all_genes <- union(filtered_genes, GENES_OF_INTEREST)
  so <- subset(raw_counts_so, cells = all_cells)
  so <- subset(so, features = all_genes)
  print("Cells with CHIKV genes of interest in the final data:")
  print(get_goi_counts(goi_cells, GENES_OF_INTEREST, so))
}

rm(raw_counts_so)
so$sample <- params$sample
so[["infected"]] <- colnames(so) %in% goi_cells
print(paste0("Created Seurat object for sample: '", params$sample,
             "' (n_cells: ", ncol(so),")"))
print(paste0("Number of infected cells: ",
             sum(so[["infected"]] == TRUE)))
```


# Compute QC measures

## Mitochondrial genes

```{r percent_mito}
mito_genes <- grep("^MT-", rownames(so@assays$RNA$counts), value = TRUE)
so <- compute_percent_mito(so, mito_genes)
```

## QC with scater

```{r qc_scater}
sce <- SingleCellExperiment(assays = list(counts = so@assays$RNA$counts))
is.mito <- grepl("^MT-", rownames(sce))
qcstats <- perCellQCMetrics(sce, subsets=list(Mito=is.mito))
filt_ad <- data.frame(
  qc.lib.low = isOutlier(qcstats$sum, log = TRUE, type = "lower"),
  qc.lib.high = isOutlier(qcstats$sum, log = FALSE, type = "higher", nmads = 4),
  qc.nexprs.low = isOutlier(qcstats$detected, log = TRUE, type = "lower"),
  qc.nexprs.high =
    isOutlier(qcstats$detected, log = FALSE, type = "higher", nmads = 4),
  qc.mito.high =
    isOutlier(qcstats$subsets_Mito_percent, log = FALSE, type = "higher")
)
qcstats <- cbind(as.data.frame(qcstats), filt_ad)
# rename sum to scater_qcstats_sum
colnames(qcstats)[colnames(qcstats) == "sum"] <- "scater_qcstats_sum"
# rename detected to scater_qcstats_detected
colnames(qcstats)[colnames(qcstats) == "detected"] <- "scater_qcstats_detected"
so@meta.data <- cbind(so@meta.data, qcstats)
```

## QC with robustbase

```{r qc_robustbase}
set.seed(745)
stats <- cbind(log10(so$nCount_RNA),
               log10(so$nFeature_RNA),
               so$percent.mito)
outlying <- adjOutlyingness(stats, only.outlyingness = TRUE)
multi.outlier <- isOutlier(outlying, type = "higher")
so$multi.outlier <- multi.outlier
```

## Removing redundant QC information

```{r rm_redundant_qc}
so@meta.data <- so@meta.data[,c("orig.ident","nCount_RNA","nFeature_RNA",
                                "scater_qcstats_sum","scater_qcstats_detected",
                                "subsets_Mito_sum","subsets_Mito_detected",
                                "percent.mito","percent.mito.cl","qc.lib.low",
                                "qc.lib.high","qc.nexprs.low","qc.nexprs.high",
                                "qc.mito.high", "multi.outlier")]
```

# Cell cycle scoring

```{r cell_cycle_scoring}
so <- test_cell_cycle_effect(so, so_name,
                             s_genes = cc.genes.updated.2019$s.genes,
                             g2m_genes = cc.genes.updated.2019$g2m.genes)
```

# Saving

```{r saving, echo=T}
saveRDS(so, paste0(var_dir, params$sample, "_so_initial.rds"))
```

# Session info

```{r session_info, echo=F}
sessionInfo()
```
