---
title: "Chikv | CellChat analysis"
author: "Giorgio Gonnella"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: sandstone
params:
  analysis: "1"
  input_step: "03.P.integration"
  input_rds: "integrated_so.rds"
  this_step: "07.P.compute_cellchat"
  output_rds: "so_w_cellchat.rds"
  condition_outer_workers: 4
  condition_ncores_per_job: 12
  sample_outer_workers: 20
  sample_ncores_per_job: 3
  overwrite_cellchat_checkpoints: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=F, warning=F, echo=F)
project_root <- here::here()
renv_activate <- file.path(project_root, "renv", "activate.R")
source(renv_activate)

worker_profile <- file.path(
  project_root,
  "analysis-1/steps/07.P.compute_cellchat/rundir/renv-worker-profile.R"
)
dir.create(dirname(worker_profile), recursive = TRUE, showWarnings = FALSE)
writeLines(sprintf("source(%s)", deparse(renv_activate)), worker_profile)
Sys.setenv(RENV_PROJECT = project_root, R_PROFILE_USER = worker_profile)

library(Seurat)
library(baia.cellchat)

options(future.globals.maxSize = 200 * 1024^3)

analysis_root <- here::here(paste0("analysis-", params$analysis,"/"))

input_step_dir <- paste0(analysis_root, "steps/", params$input_step, "/")
input_var_dir <- paste0(input_step_dir, "results/vars/")

if (!dir.exists(input_var_dir)) {
  stop("Input variable directory does not exist: ", input_var_dir)
}

this_step_dir <- paste0(analysis_root, "steps/", params$this_step, "/")
if (!dir.exists(this_step_dir)) {
  stop("Current step directory does not exist: ", this_step_dir)
}

var_dir <- paste0(this_step_dir, "results/vars/")
if (!dir.exists(var_dir)) {
  dir.create(var_dir, recursive = TRUE)
}

checkpoint_dir <- file.path(var_dir, "cellchat_checkpoints")
```

```{r load, echo=F}
so <- readRDS(file = paste0(input_var_dir, params$input_rds))
Idents(so) <- "celltype.abbrev"
```

# CellChat computation

CellChat objects are computed for each CHIKV condition and per sample. The reusable CellChat execution and derived-object construction are provided by `baia.cellchat`; this step keeps only CHIKV-specific condition and grouping configuration.

```{r cellchat_configuration, echo=TRUE}
condition_definitions <- list(
  A_NC = list(TimePoint = "A", clinical_condition = "Non-Chronic"),
  A_C = list(TimePoint = "A", clinical_condition = "Chronic"),
  `6m_NC` = list(TimePoint = "6m", clinical_condition = "Non-Chronic"),
  `6m_C` = list(TimePoint = "6m", clinical_condition = "Chronic")
)

pair_map <- list(
  A_C_vs_NC    = c("A_NC",  "A_C"),
  `6m_C_vs_NC` = c("6m_NC", "6m_C"),
  NC_A_vs_6m   = c("A_NC",  "6m_NC"),
  C_A_vs_6m    = c("A_C",   "6m_C")
)

# Cell-type grouping vector (27 cell types -> 9 groups).
group.celltype <- c(
  rep("B cells",          5),
  rep("CD4+ T cells",     8),
  rep("CD8+ T cells",     4),
  rep("GD T cells",       2),
  rep("MAIT cells",       1),
  rep("NK cells",         1),
  rep("Dendritic cells",  2),
  rep("Monocytes",        3),
  rep("Progenitor cells", 1)
)
```

```{r condition_cellchat, echo=TRUE}
so@misc$cellchat <- run_cellchat_for_conditions(
  so = so,
  condition_definitions = condition_definitions,
  group.by = "celltype.abbrev",
  ncores = params$condition_ncores_per_job,
  future_strategy = "multicore",
  outer_workers = params$condition_outer_workers,
  outer_method = "psock",
  checkpoint_dir = file.path(checkpoint_dir, "condition"),
  overwrite = params$overwrite_cellchat_checkpoints
)
```

```{r per_sample_cellchat, echo=TRUE}
so@misc$cellchat_per_sample <- run_cellchat_per_sample(
  so = so,
  condition_definitions = condition_definitions,
  group.by = "celltype.abbrev",
  sample.by = "sample",
  ncores = params$sample_ncores_per_job,
  future_strategy = "multicore",
  outer_workers = params$sample_outer_workers,
  outer_method = "psock",
  checkpoint_dir = file.path(checkpoint_dir, "per_sample"),
  overwrite = params$overwrite_cellchat_checkpoints
)
```

## Derived CellChat objects

All report visualizations read from package-style derived structures in `so@misc`.

```{r derived_cellchat_structures, echo=TRUE}
so <- add_cellchat_derived_structures(
  so = so,
  pair_definitions = pair_map,
  group.celltype = group.celltype,
  cellchat_slot = "cellchat",
  per_sample_slot = "cellchat_per_sample"
)
```

```{r save, echo=TRUE}
output_rds_path <- paste0(var_dir, params$output_rds)
if (file.exists(output_rds_path) || nzchar(Sys.readlink(output_rds_path))) {
  unlink(output_rds_path)
}
saveRDS(so, file = output_rds_path)
```
