---
title: "CHIKV | Cell proportion analyses"
author: "Giorgio Gonnella"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: sandstone
params:
  analysis: "1"
  input_step: "04.P.prop_test"
  input_rds: "so_w_prop.rds"
  step: "04.V.prop_results"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
withr::with_dir(here::here(), { source("renv/activate.R") })

library(Seurat)
library(ggplot2)
library(knitr)
library(rmdreportdeck)
library(baia.ext.cellprop.plots)

reportdeck_setup()
knitr::opts_chunk$set(cache = FALSE)

analysis_root <- here::here(paste0("analysis-", params$analysis))
input_var_dir <- file.path(analysis_root, "steps", params$input_step, "results", "vars")
this_step_dir <- file.path(analysis_root, "steps", params$step)
plots_outdir <- file.path(this_step_dir, "results", "plots")
tables_outdir <- file.path(this_step_dir, "results", "tables")

if (!dir.exists(input_var_dir)) {
  stop("Input variable directory does not exist: ", input_var_dir, call. = FALSE)
}
dir.create(plots_outdir, recursive = TRUE, showWarnings = FALSE)
dir.create(tables_outdir, recursive = TRUE, showWarnings = FALSE)

```

```{r load}
so <- readRDS(file.path(input_var_dir, params$input_rds))
```

# Goal

Test and visualize cell-type composition differences across conditions and
timepoints. Each report item contains an aligned proportion/test plot and the
canonical permutation-test table.

```{r project_configuration}
col_red  <- "#DC143C"  # Chronic, Acute
col_blue <- "#4169E1"  # Non-Chronic, 6-Months

comparison_groups_tp <- list(
  list(
    key = "A",
    data = subset(so, TimePoint == "A"),
    condition_col = "clinical_condition",
    condition_title = "Clinical Condition",
    condition_palette = c("Non-Chronic" = col_blue, "Chronic" = col_red)
  ),
  list(
    key = "6m",
    data = subset(so, TimePoint == "6m"),
    condition_col = "clinical_condition",
    condition_title = "Clinical Condition",
    condition_palette = c("Non-Chronic" = col_blue, "Chronic" = col_red)
  )
)

comparison_groups_cc <- list(
  list(
    key = "C",
    data = subset(so, clinical_condition == "Chronic"),
    condition_col = "TimePoint",
    condition_title = "Timepoint",
    condition_palette = c("A" = col_red, "6m" = col_blue)
  ),
  list(
    key = "NC",
    data = subset(so, clinical_condition == "Non-Chronic"),
    condition_col = "TimePoint",
    condition_title = "Timepoint",
    condition_palette = c("A" = col_red, "6m" = col_blue)
  )
)

cell_group_order <- sort(unique(as.character(so@meta.data$cell_group)))
cell_group_palette <- so@misc$colors$celltype_groups_bkg
cell_group_palette <- cell_group_palette[intersect(cell_group_order, names(cell_group_palette))]
```

```{r helper_functions}
item_title <- function(group, subset_name) {
  switch(
    group$key,
    A    = paste0("A timepoint, NC vs C | subsets of: ", subset_name),
    `6m` = paste0("6M timepoint, NC vs C | subsets of: ", subset_name),
    C    = paste0("C condition, A vs 6M | subsets of: ", subset_name),
    NC   = paste0("NC condition, A vs 6M | subsets of: ", subset_name),
    paste(group$key, "| subsets of:", subset_name)
  )
}

make_item_fn <- function(group) {
  function(subset_data, subset_name, test) {
    build_cellprop_report_item(
      seurat_obj    = subset_data,
      test_result   = test,
      title         = item_title(group, subset_name),
      filename_stub = paste(group$key, "singleR_proportion_test", subset_name, sep = "."),
      data_schema   = list(
        sample    = "sample",
        condition = group$condition_col,
        celltype  = "celltype.abbrev",
        group     = "cell_group"
      ),
      palette = list(
        condition = group$condition_palette,
        group     = cell_group_palette
      ),
      labels = list(
        color        = group$condition_title,
        significance = "Significance"
      ),
      layout = list(
        group_background     = TRUE,
        proportion_log_scale = TRUE
      ),
      plot_args = list(
        modifiers = list(
          combined = list(
            ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5, size = 16))
          )
        )
      )
    )
  }
}
```

# Proportion test for A timepoint: Non-Chronic vs Chronic

```{r A_results, results='asis'}
emit_cellprop_group_section(
  seurat_obj      = comparison_groups_tp[[1]]$data,
  cell_subsets    = so@misc$cell_subsets,
  test_results    = so@misc$proptest[["A"]],
  build_item_fn   = make_item_fn(comparison_groups_tp[[1]]),
  plots_dir       = plots_outdir,
  tables_dir      = tables_outdir,
  cell_group_order = cell_group_order
)
```

# Proportion test for 6M timepoint: Non-Chronic vs Chronic

```{r sixm_results, results='asis'}
emit_cellprop_group_section(
  seurat_obj      = comparison_groups_tp[[2]]$data,
  cell_subsets    = so@misc$cell_subsets,
  test_results    = so@misc$proptest[["6m"]],
  build_item_fn   = make_item_fn(comparison_groups_tp[[2]]),
  plots_dir       = plots_outdir,
  tables_dir      = tables_outdir,
  cell_group_order = cell_group_order
)
```

# Proportion test for C condition: Acute vs 6-Months

```{r C_results, results='asis'}
emit_cellprop_group_section(
  seurat_obj      = comparison_groups_cc[[1]]$data,
  cell_subsets    = so@misc$cell_subsets,
  test_results    = so@misc$proptest[["C"]],
  build_item_fn   = make_item_fn(comparison_groups_cc[[1]]),
  plots_dir       = plots_outdir,
  tables_dir      = tables_outdir,
  cell_group_order = cell_group_order
)
```

# Proportion test for NC condition: Acute vs 6-Months

```{r NC_results, results='asis'}
emit_cellprop_group_section(
  seurat_obj      = comparison_groups_cc[[2]]$data,
  cell_subsets    = so@misc$cell_subsets,
  test_results    = so@misc$proptest[["NC"]],
  build_item_fn   = make_item_fn(comparison_groups_cc[[2]]),
  plots_dir       = plots_outdir,
  tables_dir      = tables_outdir,
  cell_group_order = cell_group_order
)
```

# Session info

```{r session_info}
sessionInfo()
```
