---
title: "Chikv | CellChat results"
author: "Giorgio Gonnella"
date: 2025-12-11
output:
  html_document:
    theme: sandstone
params:
  analysis: "1"
  input_step: "07.P.compute_cellchat"
  input_rds: "so_w_cellchat.rds"
  this_step: "07.V.cellchat_results"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
withr::with_dir(here::here(), { source("renv/activate.R") })

library(Seurat)
library(CellChat)
library(patchwork)
library(ggplot2)
library(circlize)
library(ComplexHeatmap)
library(grid)

analysis_root <- here::here(paste0("analysis-", params$analysis,"/"))

input_step_dir <- paste0(analysis_root, "steps/", params$input_step, "/")
input_var_dir  <- paste0(input_step_dir, "results/vars/")

if (!dir.exists(input_var_dir)) {
  stop("Input variable directory does not exist: ", input_var_dir)
}

this_step_dir  <- paste0(analysis_root, "steps/", params$this_step, "/")
if (!dir.exists(this_step_dir)) {
  stop("Current step directory does not exist: ", this_step_dir)
}

plots_outdir <- paste0(this_step_dir, "results/plots/")
if (!dir.exists(plots_outdir)) {
  dir.create(plots_outdir, recursive = TRUE)
}
```

```{r load_so}
so <- readRDS(file = paste0(input_var_dir, params$input_rds))
```

```{r extract_cellchat_objects}
cc_A_NC  <- so@misc$cellchat$A_NC
cc_6m_NC <- so@misc$cellchat$`6m_NC`
cc_A_C <- so@misc$cellchat$A_C
cc_6m_C  <- so@misc$cellchat$`6m_C`

cellchat_list <- list(
  A_NC   = cc_A_NC,
  A_C    = cc_A_C,
  `6m_NC` = cc_6m_NC,
  `6m_C`  = cc_6m_C
)
```

# Single conditions combinations

## Overview plots

High-level exploratory plots per group: number and strength of
interactions and aggregate circle networks.

```{r plot_drawing_helper}
draw_grid_plot <- function(plt, title) {
  draw(plt, newpage=T, padding = unit(c(0.2, 0.2, 2, 0.2), "cm"))
  grid.text(title,
            y = unit(1, "npc") - unit(1, "cm"),
            gp = gpar(fontsize = 12))
} 
```


```{r overview, fig.width=7, fig.height=7}
for (nm in names(cellchat_list)) {
  cc <- cellchat_list[[nm]]
  outdir <- file.path(plots_outdir, "overview", nm)
  dir.create(outdir, showWarnings = FALSE, recursive = TRUE)

  plt <- netVisual_heatmap(cc)
  title <- paste0(nm, ": Interaction counts")
  draw_grid_plot(plt, title)
  pdf(file.path(outdir, paste0(nm, "_heatmap_count.pdf")),
      width = 7, height = 7)
  draw_grid_plot(plt, title)
  dev.off()

  plt <- netVisual_heatmap(cc, measure = "weight")
  title <- paste0(nm, ": Interaction weight")
  draw_grid_plot(plt, title)
  pdf(file.path(outdir, paste0(nm, "_heatmap_weight.pdf")),
      width = 7, height = 7)
  draw_grid_plot(plt, title)
  dev.off()
}
```

## Circle and hierarchy plots for all pathways

```{r circle_all_pathways}
for (cc_name in names(cellchat_list)) {
  cc <- cellchat_list[[cc_name]]
  all_pathways <- cc@netP$pathways
  outdir <- file.path(plots_outdir, "circle", cc_name)
  dir.create(outdir, showWarnings = F, recursive = T)
  for (pathway in all_pathways) {
    print(pathway)
    plt <- netVisual_aggregate(cc, signaling  = pathway,
      layout = "circle")
    title(paste0(cc_name, ": ", pathway))
    plt <- recordPlot()
    pdf(output_file <- file.path(outdir,
      paste0(cc_name, "_", gsub(" ", "_", pathway), "_circle.pdf")),
      width=9, height=9)
    replayPlot(plt)
    dev.off()
  }
}
```
```{r hierarchy_all_pathways}
for (cc_name in names(cellchat_list)) {
  cc <- cellchat_list[[cc_name]]
  all_pathways <- cc@netP$pathways
  outdir <- file.path(plots_outdir, "hierarchy", cc_name)
  dir.create(outdir, showWarnings = F, recursive = T)
  for (pathway in all_pathways) {
    print(pathway)
    for (cg in names(idx)) {
      print(cg)
      plt <- netVisual_aggregate(cc, signaling  = pathway,
        layout = "hierarchy", vertex.receiver = idx[[cg]],
        vertex.weight = 5, show.legend=T,
        vertex.label.cex = 0.6, point.size = 0.3,
        scale=T, big.gap=15, small.gap=2)
      title(paste0(cc_name, ": ", pathway, " - ", cg))
      plt <- recordPlot()
      pdf(output_file <- file.path(outdir,
        paste0(cc_name, "_", gsub(" ", "_", pathway),
               "_receiver_", cg, "_hierarchy.pdf")),
        width=9, height=12)
      replayPlot(plt)
      dev.off()
    }
  }
}
```

# Merged CellChat objects for binary comparisons

```{r merging_helper}
merge_two_cellchat <- function(cc1, cc2, names) {
  object.list <- setNames(list(cc1, cc2), names)
  merged      <- mergeCellChat(object.list, add.names = names(object.list))
  merged
}
```

```{r merge}
cc_A_NC_vs_C  <- merge_two_cellchat(cc_A_NC,  cc_A_C,   c("A_NC",  "A_C"))
cc_6m_NC_vs_C <- merge_two_cellchat(cc_6m_NC, cc_6m_C,  c("6m_NC", "6m_C"))
cc_NC_A_vs_6m <- merge_two_cellchat(cc_A_NC,  cc_6m_NC, c("A_NC",  "6m_NC"))
cc_C_A_vs_6m  <- merge_two_cellchat(cc_A_C,   cc_6m_C,  c("A_C",   "6m_C"))
comp_list <- list(
  A_NC_vs_C  = cc_A_NC_vs_C,
  `6m_NC_vs_C` = cc_6m_NC_vs_C,
  NC_A_vs_6m = cc_NC_A_vs_6m,
  C_A_vs_6m  = cc_C_A_vs_6m
)
```

```{r, fig.width=12, fig.height=12}
ggA1 <- rankNet(cc_A_NC_vs_C,  mode="comparison", measure="weight", stacked=TRUE,  do.stat=FALSE)
ggA2 <- rankNet(cc_A_NC_vs_C,  mode="comparison", measure="weight", stacked=FALSE, do.stat=FALSE)
gg6_1 <- rankNet(cc_6m_NC_vs_C, mode="comparison", measure="weight", stacked=TRUE,  do.stat=FALSE)
gg6_2 <- rankNet(cc_6m_NC_vs_C, mode="comparison", measure="weight", stacked=FALSE, do.stat=FALSE)

(ggA1 + ggA2) / (gg6_1 + gg6_2)
```


```{r, fig.width=12, fig.height=9}
library(CellChat)
library(ComplexHeatmap)
library(grid)  # for unit()

# Helper: make two comparable heatmaps (same pathway set) side-by-side
make_pair_ht <- function(cc_left, cc_right, left_title, right_title,
                         pattern = "outgoing",
                         width = 5, height = 6, gap_cm = 0.5) {
  
  # Union of pathways to ensure same rows appear in both heatmaps
  pathway.union <- union(cc_left@netP$pathways, cc_right@netP$pathways)

  ht_left <- netAnalysis_signalingRole_heatmap(
    cc_left,
    pattern   = pattern,
    signaling = pathway.union,
    title     = left_title,
    width     = width,
    height    = height
  )

  ht_right <- netAnalysis_signalingRole_heatmap(
    cc_right,
    pattern   = pattern,
    signaling = pathway.union,
    title     = right_title,
    width     = width,
    height    = height
  )

  draw(ht_left + ht_right, ht_gap = unit(gap_cm, "cm"))
  
  invisible(list(ht_left = ht_left, ht_right = ht_right, pathways = pathway.union))
}

# Panels C–D
# Acute: A_C and A_NC side-by-side
acute_ht <- make_pair_ht(
  cc_left    = cc_A_NC,
  cc_right   = cc_A_C,
  left_title = "A_NC",
  right_title= "A_C",
  pattern    = "outgoing",
  width      = 10, height = 12, gap_cm = 1.5
)

# 6 months: 6m_C and 6m_NC side-by-side
m6_ht <- make_pair_ht(
  cc_left    = cc_6m_NC,
  cc_right   = cc_6m_C,
  left_title = "6m_NC",
  right_title= "6m_C",
  pattern    = "outgoing",
  width      = 10, height = 12, gap_cm = 1.5
)

```

```{r merged, fig.width=12, fig.height=10}
for (cc_name in names(comp_list)) {
  cc <- comp_list[[cc_name]]
  outdir <- file.path(plots_outdir, "comparisons", cc_name)
  dir.create(outdir, showWarnings = F, recursive = T)
  # # Comparing the total number of interactions
  gg1 <- compareInteractions(cc, show.legend = F, group = c(1,2)) +
     ggtitle(paste0(cc_name, ": Number of interactions"))
  print(gg1)
  ggsave(filename = file.path(outdir,
         paste0(cc_name, "_compareInteractions_count.pdf")),
         plot = gg1, width = 7, height = 5)
  # Comparing the total interaction strength
  gg2 <- compareInteractions(cc, show.legend = F, group = c(1,2),
                             measure="weight") +
  ggtitle(paste0(cc_name, ": Interaction strength"))
  print(gg2)
  ggsave(filename = file.path(outdir,
         paste0(cc_name, "_compareInteractions_weight.pdf")),
    plot = gg2, width = 7, height = 5)
  # Examine the differential number of interactions
  op <- par(mar = c(2, 2, 6, 2), xpd = NA)
  netVisual_diffInteraction(cc, weight.scale = T)
  title(paste0(cc_name, ": Differential number of interactions"), cex.main = 1.4, line = 2)
  grDevices::pdf(file.path(outdir,
    paste0(cc_name, "_diffInteraction_count.pdf")),
    width = 12, height = 10)
  dev.off()
  # Examine the differential interaction strengths
  op <- par(mar = c(2, 2, 6, 2), xpd = NA)
  netVisual_diffInteraction(cc, weight.scale = T, measure = "weight")
  title(paste0(cc_name, ": Differential interaction strengths"), cex.main = 1.4, line = 2)
  grDevices::pdf(file.path(outdir,
    paste0(cc_name, "_diffInteraction_weight.pdf")),
    width = 12, height = 10)
  dev.off()
  # heatmap of differential number of interactions
  plt <- netVisual_heatmap(cc)
  title <- paste0(cc_name, ": Differential interaction counts")
  draw_grid_plot(plt, title)
  pdf(file.path(outdir, paste0(cc_name, "_diff_heatmap_count.pdf")),
      width = 7, height = 7)
  draw_grid_plot(plt, title)
  dev.off()

  # heatmap of differential interaction strengths
  plt <- netVisual_heatmap(cc, measure = "weight")
  title <- paste0(cc_name, ": Differential interaction weights")
  draw_grid_plot(plt, title)
  pdf(file.path(outdir, paste0(cc_name, "_diff_heatmap_weight.pdf")),
      width = 7, height = 7)
  draw_grid_plot(plt, title)
  dev.off()
}
```

# Cell groups

```{r index_cell_groups}
idx <- list()
idx[["B_cells"]]           <- c(1,2,3,4,5)
idx[["CD4_T_cells"]]       <- c(6,7,8,9,10,11,12,13)
idx[["CD8_T_cells"]]       <- c(14,15,16,17)
idx[["GD_T_cells"]]        <- c(18,19)
idx[["Dendritic_cells"]]   <- c(22,23)
idx[["Monocytes"]]         <- c(24,25,26)
idx[["Other_cells"]]       <- c(20, 21, 27)

group.celltype <- c(
  rep("B cells",          5),
  rep("CD4+ T cells",     8),
  rep("CD8+ T cells",     4),
  rep("GD T cells",       2),
  rep("MAIT cells",       1),
  rep("NK cells",         1),
  rep("Dendritic cells",  2),
  rep("Monocytes",        3),
  rep("Progenitor cells", 1)
)
```

```{r merge_interactions}
mergeInteractions_inplace <- function(cc, group.celltype) {
  cc_g <- mergeInteractions(cc, group.celltype)
  cc_g@net$count  <- cc_g@net$count.merged
  cc_g@net$weight <- cc_g@net$weight.merged
  cc_g@net$count.merged  <- NULL
  cc_g@net$weight.merged <- NULL
  cc_g@idents <- factor(group.celltype[as.integer(cc@idents)],
                      levels = unique(group.celltype))
  return(cc_g)
}

cc_A_NC_g  = mergeInteractions_inplace(cc_A_NC, group.celltype)
cc_6m_NC_g = mergeInteractions_inplace(cc_6m_NC, group.celltype)
cc_A_C_g   = mergeInteractions_inplace(cc_A_C, group.celltype)
cc_6m_C_g  = mergeInteractions_inplace(cc_6m_C, group.celltype)

grouped_list <- list(
  A_NC   = cc_A_NC_g,
  A_C    = cc_A_C_g,
  `6m_NC` = cc_6m_NC_g,
  `6m_C`  = cc_6m_C_g
)
```

# Single grouped CellChat objects

```{r grouped_cellchat_plots}
for (nm in names(grouped_list)) {
  cc <- grouped_list[[nm]]
  
  outdir <- file.path(plots_outdir, "overview_grouped", nm)
  dir.create(outdir, showWarnings = FALSE, recursive = TRUE)
  
  # Heatmap of grouped interaction counts
  plt <- netVisual_heatmap(cc)
  title <- paste0(nm, ": Interaction counts")
  draw_grid_plot(plt, title)
  pdf(file.path(outdir, paste0(nm, "_grouped_heatmap_count.pdf")),
      width = 7, height = 7)
  draw_grid_plot(plt, title)
  dev.off()

  # Heatmap of grouped interaction strengths
  plt <- netVisual_heatmap(cc, measure = "weight")
  title <- paste0(nm, ": Interaction weight")
  draw_grid_plot(plt, title)
  pdf(file.path(outdir, paste0(nm, "_grouped_heatmap_weight.pdf")),
      width = 7, height = 7)
  draw_grid_plot(plt, title)
  dev.off()
}
```

## Merged grouped CellChat objects for binary comparisons

```{r merge_cell_groups}
cc_NC_A_vs_6m_g <- merge_two_cellchat(cc_A_NC_g, cc_6m_NC_g,c("A_NC", "6m_NC"))
cc_C_A_vs_6m_g  <- merge_two_cellchat(cc_A_C_g,  cc_6m_C_g, c("A_C",  "6m_C"))
cc_A_NC_vs_C_g  <- merge_two_cellchat(cc_A_NC_g, cc_A_C_g,  c("A_NC", "A_C"))
cc_6m_NC_vs_C_g <- merge_two_cellchat(cc_6m_NC_g,cc_6m_C_g, c("6m_NC","6m_C"))

comp_grouped_list <- list(
  NC_A_vs_6m = cc_NC_A_vs_6m_g,
  C_A_vs_6m  = cc_C_A_vs_6m_g,
  A_NC_vs_C  = cc_A_NC_vs_C_g,
  `6m_NC_vs_C` = cc_6m_NC_vs_C_g
)
```

```{r merged, fig.width=12, fig.height=10}
for (cc_name in names(comp_grouped_list)) {
  cc <- comp_grouped_list[[cc_name]]
  outdir <- file.path(plots_outdir, "comparisons_grouped", cc_name)
  dir.create(outdir, showWarnings = F, recursive = T)
  # # Comparing the total number of interactions
  gg1 <- compareInteractions(cc, show.legend = F, group = c(1,2)) +
     ggtitle(paste0(cc_name, ": Number of interactions"))
  print(gg1)
  ggsave(filename = file.path(outdir,
         paste0(cc_name, "_grouped_compareInteractions_count.pdf")),
         plot = gg1, width = 7, height = 5)
  # Comparing the total interaction strength
  gg2 <- compareInteractions(cc, show.legend = F, group = c(1,2),
                             measure="weight") +
  ggtitle(paste0(cc_name, ": Interaction strength"))
  print(gg2)
  ggsave(filename = file.path(outdir,
         paste0(cc_name, "_grouped_compareInteractions_weight.pdf")),
    plot = gg2, width = 7, height = 5)
  # Examine the differential number of interactions
  op <- par(mar = c(2, 2, 6, 2), xpd = NA)
  netVisual_diffInteraction(cc, weight.scale = T)
  title(paste0(cc_name, ": Differential number of interactions"), cex.main = 1.4, line = 2)
  grDevices::pdf(file.path(outdir,
    paste0(cc_name, "_grouped_diffInteraction_count.pdf")),
    width = 12, height = 10)
  dev.off()
  # Examine the differential interaction strengths
  op <- par(mar = c(2, 2, 6, 2), xpd = NA)
  netVisual_diffInteraction(cc, weight.scale = T, measure = "weight")
  title(paste0(cc_name, ": Differential interaction strengths"), cex.main = 1.4, line = 2)
  grDevices::pdf(file.path(outdir,
    paste0(cc_name, "_grouped_diffInteraction_weight.pdf")),
    width = 12, height = 10)
  dev.off()
  # heatmap of differential number of interactions
  plt <- netVisual_heatmap(cc)
  title <- paste0(cc_name, ": Differential interaction counts")
  draw_grid_plot(plt, title)
  pdf(file.path(outdir, paste0(cc_name, "_grouped_diff_heatmap_count.pdf")),
      width = 7, height = 7)
  draw_grid_plot(plt, title)
  dev.off()

  # heatmap of differential interaction strengths
  plt <- netVisual_heatmap(cc, measure = "weight")
  title <- paste0(cc_name, ": Differential interaction weights")
  draw_grid_plot(plt, title)
  pdf(file.path(outdir, paste0(cc_name, "_grouped_diff_heatmap_weight.pdf")),
      width = 7, height = 7)
  draw_grid_plot(plt, title)
  dev.off()
}
```

## Session info

```{r, echo=F}
sessionInfo()
```
