---
title: "`r paste('Chikv | CellChat', switch(params$pair, A_C_vs_NC='Acute - Chronic vs Non-Chronic', '6m_C_vs_NC'='6 months - Chronic vs Non-Chronic', NC_A_vs_6m='Non-Chronic - Acute vs 6 months', C_A_vs_6m='Chronic - Acute vs 6 months'), c(all='unfiltered', pfilt='p-value filtered')[params$sfx], sep=' | ')`"
author: "Giorgio Gonnella"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: sandstone
params:
  analysis: "1"
  input_step: "07.P.compute_cellchat"
  input_rds: "so_w_cellchat.rds"
  this_step: "07.V.cellchat_results"
  pair: "A_C_vs_NC"
  sfx:  "pfilt"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
withr::with_dir(here::here(), { source("renv/activate.R") })

library(Seurat)
library(CellChat)
library(patchwork)
library(ggplot2)
library(circlize)
library(ComplexHeatmap)
library(grid)
library(rmdreportdeck)
reportdeck_setup()
knitr::opts_chunk$set(cache=F)

analysis_root <- here::here(paste0("analysis-", params$analysis,"/"))
input_var_dir <- paste0(analysis_root, "steps/", params$input_step, "/results/vars/")
if (!dir.exists(input_var_dir))
  stop("Input variable directory does not exist: ", input_var_dir)
this_step_dir <- paste0(analysis_root, "steps/", params$this_step, "/")
if (!dir.exists(this_step_dir))
  stop("Current step directory does not exist: ", this_step_dir)

pair_map <- list(
  A_C_vs_NC    = c("A_NC",  "A_C"),
  `6m_C_vs_NC` = c("6m_NC", "6m_C"),
  NC_A_vs_6m   = c("A_NC",  "6m_NC"),
  C_A_vs_6m    = c("A_C",   "6m_C")
)
if (!params$pair %in% names(pair_map))
  stop("Unknown pair: ", params$pair)
if (!params$sfx %in% c("all", "pfilt"))
  stop("Unknown sfx: ", params$sfx)

conds  <- pair_map[[params$pair]]
nm1    <- conds[1]; nm2 <- conds[2]
sfx    <- params$sfx
thresh <- if (sfx == "pfilt") 0.05 else 1
pfx    <- paste0(params$pair, ".", sfx)

plots_outdir <- paste0(this_step_dir, "results/", params$pair, "/", sfx, "/plots/")
dir.create(plots_outdir, recursive = TRUE, showWarnings = FALSE)

report_href <- function(path) {
  file.path("..", "results", params$pair, sfx, "plots", path)
}

.fig_counter <- 0L
fig_label <- function(desc = NULL, png_href = NULL, png_fname = NULL,
                      pdf_href = NULL, pdf_fname = NULL,
                      tsv_href = NULL, tsv_fname = NULL) {
  .fig_counter <<- .fig_counter + 1L
  figure_id <- paste0(.fig_counter, ".", pfx)
  lbl <- paste0(figure_id, if (!is.null(desc)) paste0(" \u2014 ", desc) else "")
  cat(paste0("\n\n### ", lbl, "\n\n"))
  dl <- list()
  if (!is.null(png_href))
    dl <- c(dl, list(report_download_link("Download PNG",
      if (is.null(png_fname)) basename(png_href) else png_fname, png_href)))
  if (!is.null(pdf_href))
    dl <- c(dl, list(report_download_link("Download PDF",
      if (is.null(pdf_fname)) basename(pdf_href) else pdf_fname, pdf_href)))
  if (!is.null(tsv_href))
    dl <- c(dl, list(report_download_link("Download TSV",
      if (is.null(tsv_fname)) basename(tsv_href) else tsv_fname, tsv_href)))
  cat(as.character(do.call(report_asset_bar,
    c(list(report_figure_name(figure_id)), dl))), "\n")
  invisible(figure_id)
}
```

# Goal

Summarize CellChat cell–cell communication results for a given condition pair and filter setting.
Displays overview heatmaps, interaction count bar charts, signaling role heatmaps, chord diagrams,
rank-net plots, and bubble plots comparing signaling between conditions, with statistical tests
on per-sample values where applicable.

```{r helpers}
draw_grid_plot <- function(plt, title) {
  draw(plt, newpage = TRUE, padding = unit(c(0.2, 0.2, 2, 0.2), "cm"))
  grid.text(title, y = unit(1, "npc") - unit(1, "cm"),
            gp = gpar(fontsize = 12))
}

make_pair_ht <- function(cc_left, cc_right, left_title, right_title,
                         pattern = "outgoing", width = 5, height = 6, gap_cm = 0.5,
                         pdf_path = NULL, png_path = NULL) {
  pathway.union <- union(cc_left@netP$pathways, cc_right@netP$pathways)
  ht_left  <- netAnalysis_signalingRole_heatmap(cc_left,  pattern = pattern,
    signaling = pathway.union, title = left_title,  width = width, height = height)
  ht_right <- netAnalysis_signalingRole_heatmap(cc_right, pattern = pattern,
    signaling = pathway.union, title = right_title, width = width, height = height)
  if (!is.null(png_path)) {
    png(png_path, width = (width * 2 + gap_cm + 2) * 100, height = (height + 2) * 100)
    draw(ht_left + ht_right, ht_gap = unit(gap_cm, "cm"))
    invisible(dev.off())
  }
  if (!is.null(pdf_path)) {
    pdf(pdf_path, width = (width * 2 + gap_cm) + 2, height = height + 2)
    draw(ht_left + ht_right, ht_gap = unit(gap_cm, "cm"))
    invisible(dev.off())
  }
  draw(ht_left + ht_right, ht_gap = unit(gap_cm, "cm"))
}

# --- Note helpers ---
# Filtering method used for pfilt:
#   CellChat permutation test (100 permutations per LR pair per sender-receiver pair),
#   p <= 0.05 on communication probability.
# Between-condition statistical test:
#   Wilcoxon rank-sum on per-sample values (two-sided, normal approximation),
#   BH correction for multiple testing where applied.

FILT_METHOD <- paste0("CellChat permutation test on communication probability",
                      " (100 permutations per ligand\u2013receptor pair per sender\u2013receiver pair)")

write_tsv <- function(x, path) {
  dir.create(dirname(path), recursive = TRUE, showWarnings = FALSE)
  write.table(x, file = path, sep = "\t", quote = FALSE, row.names = FALSE, na = "")
  invisible(path)
}

matrix_to_long <- function(m, value_name = "value") {
  out <- expand.grid(source = rownames(m), target = colnames(m), stringsAsFactors = FALSE)
  out[[value_name]] <- as.vector(m)
  out
}

interaction_matrix_df <- function(cc_obj, condition, measure = "count") {
  value_name <- if (measure == "count") "lr_interaction_count" else "communication_probability"
  df <- matrix_to_long(if (measure == "count") cc_obj@net$count else cc_obj@net$weight,
                       value_name = value_name)
  df$condition <- condition
  df[, c("condition", "source", "target", value_name)]
}

pathway_matrix_df <- function(cc_obj, condition, pathway) {
  lr_names <- dimnames(cc_obj@net$prob)[[3]]
  pw_map <- cc_obj@LR$LRsig[lr_names, "pathway_name"]
  lr_idx <- which(pw_map == pathway)
  prob <- cc_obj@netP$prob[, , pathway, drop = TRUE]
  count <- if (length(lr_idx) == 0) prob * 0 else
    apply(cc_obj@net$prob[, , lr_idx, drop = FALSE] > 0, c(1, 2), sum)
  df <- matrix_to_long(prob, value_name = "communication_probability")
  df$lr_interaction_count <- as.vector(count)
  df$condition <- condition
  df$pathway <- pathway
  df[, c("condition", "pathway", "source", "target",
         "lr_interaction_count", "communication_probability")]
}

signaling_role_df <- function(cc_obj, condition, pattern = "outgoing") {
  cent <- cc_obj@netP$centr
  if (is.null(cent)) return(data.frame())
  rows <- lapply(names(cent), function(pathway) {
    vals <- cent[[pathway]][[pattern]]
    if (is.null(vals)) return(NULL)
    data.frame(
      condition = condition,
      pathway = pathway,
      cell_type = names(vals),
      score = as.numeric(vals),
      stringsAsFactors = FALSE
    )
  })
  df <- do.call(rbind, rows)
  if (is.null(df)) return(data.frame())
  names(df)[names(df) == "score"] <- paste0(pattern, "_score")
  df
}

merged_diff_df <- function(merged_obj, measure = "count") {
  value_name <- if (measure == "count") "lr_interaction_count_difference" else
    "communication_probability_difference"
  mats <- if (measure %in% names(merged_obj@net)) merged_obj@net[[measure]] else
    list(merged_obj@net[[nm1]][[measure]], merged_obj@net[[nm2]][[measure]])
  diff <- if (is.list(mats) && length(mats) >= 2) mats[[1]] - mats[[2]] else mats
  df <- matrix_to_long(diff, value_name = value_name)
  df$comparison <- paste0(nm1, "_minus_", nm2)
  df[, c("comparison", "source", "target", value_name)]
}

gg_data_df <- function(gg, panel = NULL) {
  df <- gg$data
  if (is.null(df)) return(data.frame())
  if (!is.null(panel)) df$panel <- panel
  df
}

# --- Per-sample stats infrastructure ---

# group.celltype.vec: maps positions 1:27 (cell types in level order) to 9 cell groups.
# Must match the group.celltype vector used in compute_cellchat.Rmd.
group.celltype.vec <- c(
  rep("B cells",          5),
  rep("CD4+ T cells",     8),
  rep("CD8+ T cells",     4),
  rep("GD T cells",       2),
  rep("MAIT cells",       1),
  rep("NK cells",         1),
  rep("Dendritic cells",  2),
  rep("Monocytes",        3),
  rep("Progenitor cells", 1)
)

# Extract per-sample net matrices, subset to cell types common across all samples.
# Returns NULL if <2 samples per group.
.persample_mats <- function(measure = "count") {
  ps1 <- so@misc[["cellchat_per_sample"]][[nm1]]
  ps2 <- so@misc[["cellchat_per_sample"]][[nm2]]
  if (is.null(ps1) || is.null(ps2) || length(ps1) < 2 || length(ps2) < 2)
    return(NULL)
  all_ps      <- c(ps1, ps2)
  common_cts  <- Reduce(intersect, lapply(all_ps, function(cc_s) levels(cc_s@idents)))
  if (length(common_cts) < 2) return(NULL)
  # For pfilt, mask per-sample matrices to only include sender–receiver pairs
  # that are significant at the condition level (non-zero in the pfilt cc object).
  make_mask <- function(cc_cond) {
    idx <- match(common_cts, levels(cc_cond@idents))
    if (any(is.na(idx))) return(NULL)
    m <- if (measure == "count") cc_cond@net$count else cc_cond@net$weight
    m[idx, idx, drop = FALSE] > 0
  }
  mask1 <- if (sfx == "pfilt") make_mask(cc[[nm1]]) else NULL
  mask2 <- if (sfx == "pfilt") make_mask(cc[[nm2]]) else NULL
  get_mat <- function(cc_s, mask) {
    idx <- match(common_cts, levels(cc_s@idents))
    m   <- if (measure == "count") cc_s@net$count else cc_s@net$weight
    m   <- structure(m[idx, idx, drop = FALSE], dimnames = list(common_cts, common_cts))
    if (!is.null(mask)) m[!mask] <- 0
    m
  }
  list(mats1 = lapply(ps1, get_mat, mask = mask1),
       mats2 = lapply(ps2, get_mat, mask = mask2),
       cts = common_cts, n1 = length(ps1), n2 = length(ps2))
}

# Aggregate per-sample ungrouped matrices to the 9 cell-group level.
.persample_grouped_mats <- function(measure = "count") {
  d <- .persample_mats(measure)
  if (is.null(d)) return(NULL)
  ref_lev  <- levels(so@misc$cc$all[[nm1]]@idents)   # 27 full reference levels
  pos      <- match(d$cts, ref_lev)
  grps     <- group.celltype.vec[pos]
  uniq_grps <- unique(group.celltype.vec)             # 9 groups in order
  agg <- function(m) {
    out <- matrix(0, length(uniq_grps), length(uniq_grps),
                  dimnames = list(uniq_grps, uniq_grps))
    for (i in seq_along(grps))
      for (j in seq_along(grps))
        out[grps[i], grps[j]] <- out[grps[i], grps[j]] + m[i, j]
    out
  }
  list(mats1 = lapply(d$mats1, agg), mats2 = lapply(d$mats2, agg),
       cts = uniq_grps, n1 = d$n1, n2 = d$n2)
}

.fmt_p <- function(p) if (is.null(p) || is.na(p)) "NA" else if (p < 0.001) sprintf("%.2e", p) else sprintf("%.3f", p)
.stars <- function(p) if (is.null(p) || is.na(p)) "" else if (p < 0.001) "***" else if (p < 0.01) "**" else if (p < 0.05) "*" else "ns"

# Wilcoxon rank-sum on per-sample totals (for Fig8-9, Fig18-19).
.wilcox_total <- function(measure = "count") {
  d <- .persample_mats(measure)
  if (is.null(d)) return(NULL)
  t1 <- sapply(d$mats1, sum)
  t2 <- sapply(d$mats2, sum)
  wt <- suppressWarnings(wilcox.test(t1, t2, exact = FALSE))
  list(stat = wt$statistic, p = wt$p.value, n1 = d$n1, n2 = d$n2,
       med1 = median(t1), med2 = median(t2))
}

# Wilcoxon rank-sum per sender-receiver pair with BH correction (for Fig10-13, Fig20-23).
.wilcox_perpair <- function(measure = "count", is_grouped = FALSE) {
  d <- if (is_grouped) .persample_grouped_mats(measure) else .persample_mats(measure)
  if (is.null(d)) return(NULL)
  n    <- length(d$cts)
  pmat <- matrix(NA_real_, n, n, dimnames = list(d$cts, d$cts))
  for (i in seq_len(n)) for (j in seq_len(n)) {
    v1 <- sapply(d$mats1, function(m) m[i, j])
    v2 <- sapply(d$mats2, function(m) m[i, j])
    if (any(c(v1, v2) != 0))
      pmat[i, j] <- tryCatch(
        suppressWarnings(wilcox.test(v1, v2, exact = FALSE)$p.value),
        error = function(e) NA_real_)
  }
  tested <- !is.na(pmat)
  pmat_adj <- pmat
  if (any(tested))
    pmat_adj[tested] <- p.adjust(pmat[tested], method = "BH")
  list(p.adj = pmat_adj, cts = d$cts, n1 = d$n1, n2 = d$n2,
       n_tested = sum(tested))
}

# Summarize CellChat rankNet(do.stat=TRUE) pathway p-values.
.ranknet_stats <- function(ranknet_df, thresh = 0.05, p_adjust_method = "BH") {
  if (is.null(ranknet_df) || !"pvalues" %in% names(ranknet_df))
    return(NULL)
  p_by_pw <- tapply(ranknet_df$pvalues, as.character(ranknet_df$name), function(x) {
    x <- x[!is.na(x)]
    if (length(x) == 0) NA_real_ else x[1]
  })
  tested <- !is.na(p_by_pw)
  p_adj <- rep(NA_real_, length(p_by_pw))
  names(p_adj) <- names(p_by_pw)
  p_adj[tested] <- p.adjust(p_by_pw[tested], method = p_adjust_method)
  list(p_raw = p_by_pw,
       p_adj = p_adj,
       p_adjust_method = p_adjust_method,
       n_tested = sum(tested),
       n_sig = sum(p_adj <= thresh, na.rm = TRUE))
}

# Post-process rankNet ggplot: recolor y-axis (pathway) labels by significance.
# gg        — ggplot returned by rankNet
# pw_stats  — result of .ranknet_stats()
# thresh    — significance threshold (default 0.05)
# Colors:
#   red  (#E41A1C) = significantly stronger in nm1 (rankNet FDR < thresh)
#   blue (#377EB8) = significantly stronger in nm2 (rankNet FDR < thresh)
#   black = not significant
.color_ranknet_labels <- function(gg, pw_stats, thresh = 0.05) {
  if (is.null(pw_stats)) return(gg)
  pvals <- pw_stats$p_adj
  gg + theme(axis.text.y = ggplot2::element_text(
    colour = sapply(levels(gg$data$name %||% factor(gg$data$name)), function(pw) {
      p <- pvals[pw]
      if (is.na(p) || p > thresh) return("black")
      # Determine direction from data using nm1/nm2 directly (not level order,
      # which may differ from nm1/nm2 order in the merged CellChat object).
      d <- gg$data[gg$data$name == pw, ]
      if (nrow(d) == 0) return("black")
      g1 <- sum(d$contribution[d$group == nm1], na.rm=TRUE)
      g2 <- sum(d$contribution[d$group == nm2], na.rm=TRUE)
      if (g1 >= g2) "#E41A1C" else "#377EB8"  # red=nm1 stronger, blue=nm2
    })
  ))
}

.add_ranknet_pvalues <- function(gg, pw_stats) {
  if (is.null(pw_stats) || is.null(gg$data) || !"name" %in% names(gg$data))
    return(gg)
  pw <- as.character(gg$data$name)
  gg$data$pvalues_raw <- pw_stats$p_raw[pw]
  gg$data$pvalues_BH <- pw_stats$p_adj[pw]
  gg
}

# Formatted note for the total Wilcoxon test result.
.total_test_note <- function(measure = "count") {
  wt <- .wilcox_total(measure)
  if (is.null(wt)) {
    cat("Between-condition test: not available (need \u22652 samples per group).\n\n")
  } else {
    unit <- if (measure == "count") "total ligand\u2013receptor (LR) interaction count" else "total communication probability"
    sig  <- wt$p < 0.05
    sig_str <- if (sig) {
      paste0("statistically significant (", .stars(wt$p), ")")
    } else {
      "not statistically significant"
    }
    cat(paste0(
      "Between-condition test (", nm1, " vs ", nm2, "): ",
      "Wilcoxon rank-sum on per-sample ", unit,
      " (n\u2081=", wt$n1, ", n\u2082=", wt$n2, " samples; ",
      "W=", wt$stat, ", p=", .fmt_p(wt$p),
      "; median ", round(wt$med1, 1), " vs ", round(wt$med2, 1), "). ",
      "The observed difference between the two conditions is ",
      sig_str, " according to this test.\n\n"
    ))
  }
}

# Per-sample barplot: one bar per sample, colored by condition.
# colors: named vector c(nm1 = "#hex", nm2 = "#hex") — should match the
#   compareInteractions plot; pass the result of .extract_ci_colors(gg).
# Draws and prints the ggplot, returns it invisibly.
.extract_ci_colors <- function(gg) {
  built <- ggplot_build(gg)$data[[1]]
  grp_fill <- unique(built[c("group", "fill")])
  grp_fill <- grp_fill[order(grp_fill$group), ]
  setNames(grp_fill$fill, c(nm1, nm2))
}

.persample_barplot <- function(measure = "count", is_grouped = FALSE,
                               colors = NULL, tsv_path = NULL) {
  d <- if (is_grouped) .persample_grouped_mats(measure) else .persample_mats(measure)
  if (is.null(d)) {
    cat("Per-sample barplot: not available (need \u22652 samples per group).\n\n")
    return(invisible(NULL))
  }
  unit  <- if (measure == "count") "LR interaction count" else "communication probability"
  vals1 <- sapply(d$mats1, sum)
  vals2 <- sapply(d$mats2, sum)
  vals_all <- NULL
  if (sfx == "pfilt") {
    old_sfx <- sfx
    sfx <<- "all"
    d_all <- if (is_grouped) .persample_grouped_mats(measure) else .persample_mats(measure)
    sfx <<- old_sfx
    if (!is.null(d_all)) {
      vals_all <- c(sapply(d_all$mats1, sum), sapply(d_all$mats2, sum))
    }
  }
  df <- data.frame(
    value     = c(vals1, vals2),
    sample    = factor(c(names(d$mats1), names(d$mats2)),
                       levels = c(names(d$mats1), names(d$mats2))),
    condition = factor(c(rep(nm1, d$n1), rep(nm2, d$n2)), levels = c(nm1, nm2))
  )
  if (!is.null(vals_all) && length(vals_all) == nrow(df)) {
    df$unfiltered_value <- vals_all
    df$filtered_value <- df$value
    df$removed_by_filter <- df$unfiltered_value - df$filtered_value
  }
  if (!is.null(tsv_path)) write_tsv(df, tsv_path)
  wt <- suppressWarnings(wilcox.test(vals1, vals2, exact = FALSE))
  p_label <- paste0("Wilcoxon p = ", .fmt_p(wt$p.value))
  gg <- ggplot(df, aes(x = sample, y = value, fill = condition)) +
    geom_col(width = 0.7) +
    labs(x = NULL, y = unit, fill = "Condition",
         title = paste0(params$pair, " (", sfx, "): ", unit, " per sample"),
         caption = p_label) +
    theme_bw(base_size = 11) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  if (!is.null(colors))
    gg <- gg + scale_fill_manual(values = colors)
  print(gg)
  filt_note <- if (sfx == "pfilt" && "removed_by_filter" %in% names(df)) {
    removed <- paste0(as.character(df$sample), ": ", df$removed_by_filter,
                      " removed (", df$filtered_value, "/", df$unfiltered_value,
                      " retained)", collapse = "; ")
    paste0(" Only sender\u2013receiver pairs significant at the condition level",
           " (CellChat permutation test, p\u00a0\u22640.05) are included. ",
           "Filtered interactions per sample: ", removed, ".")
  } else if (sfx == "pfilt") {
    paste0(" Only sender\u2013receiver pairs significant at the condition level",
           " (CellChat permutation test, p\u00a0\u22640.05) are included.")
  } else ""
  cat(paste0(
    "Total ", unit, " for each individual sample, colored by condition",
    " (see legend). ",
    "The spread of bars within each condition shows within-group variability. ",
    "Between-condition test: Wilcoxon rank-sum, p = ", .fmt_p(wt$p.value), ".",
    filt_note, "\n\n"
  ))
  invisible(gg)
}

# Per-pair test note (text only, for chord diagrams Fig10-11, Fig20-21).
.perpair_chord_note <- function(measure = "count", is_grouped = FALSE) {
  wp <- .wilcox_perpair(measure, is_grouped)
  pair_type <- if (is_grouped) "cell-group" else "cell-type"
  if (is.null(wp)) {
    cat(paste0("*Per-", pair_type, "-pair test: not available",
               " (need \u22652 samples per group).*\n\n"))
    return(invisible(NULL))
  }
  pmat   <- wp$p.adj
  n_sig  <- sum(pmat <= 0.05, na.rm = TRUE)
  if (n_sig == 0) {
    cat(paste0(
      "*Per-", pair_type, "-pair Wilcoxon rank-sum with BH correction ",
      "(n\u2081=", wp$n1, ", n\u2082=", wp$n2, " samples, ",
      wp$n_tested, " pairs tested): no significant pairs (p.adj\u00a0>\u00a00.05).*\n\n"
    ))
  } else {
    sig_idx  <- which(pmat <= 0.05, arr.ind = TRUE)
    pair_str <- apply(sig_idx, 1, function(r)
      paste0(rownames(pmat)[r[1]], "\u2192", colnames(pmat)[r[2]],
             "\u00a0", .stars(pmat[r[1], r[2]])))
    cat(paste0(
      "*Per-", pair_type, "-pair Wilcoxon rank-sum with BH correction ",
      "(n\u2081=", wp$n1, ", n\u2082=", wp$n2, " samples, ",
      wp$n_tested, " pairs tested): ",
      n_sig, " significant pair(s): ",
      paste(pair_str, collapse = "; "), ".*\n\n"
    ))
  }
}

# Significance heatmap figure for differential heatmaps (Fig12-13, Fig22-23).
# Draws a -log10(p.adj) heatmap with asterisk overlay and returns text note.
.sig_heatmap_fig <- function(measure = "count", is_grouped = FALSE) {
  wp <- .wilcox_perpair(measure, is_grouped)
  pair_type <- if (is_grouped) "cell-group" else "cell-type"
  if (is.null(wp)) {
    cat(paste0("*Per-", pair_type, "-pair significance heatmap: not available",
               " (need \u22652 samples per group).*\n\n"))
    return(invisible(NULL))
  }
  pmat  <- wp$p.adj
  n_sig <- sum(pmat <= 0.05, na.rm = TRUE)
  # Text note
  cat(paste0(
    "*Per-", pair_type, "-pair Wilcoxon rank-sum with BH correction ",
    "(n\u2081=", wp$n1, ", n\u2082=", wp$n2, " samples, ",
    wp$n_tested, " pairs tested): ",
    n_sig, "/", wp$n_tested, " significant at p.adj\u00a0\u22640.05.*\n\n"
  ))
  # Build significance heatmap
  logp      <- matrix(0, nrow(pmat), ncol(pmat), dimnames = dimnames(pmat))
  logp[!is.na(pmat)] <- -log10(pmax(pmat[!is.na(pmat)], 1e-300))
  stars_mat <- matrix("", nrow(pmat), ncol(pmat), dimnames = dimnames(pmat))
  for (i in seq_len(nrow(pmat))) for (j in seq_len(ncol(pmat)))
    if (!is.na(pmat[i, j])) stars_mat[i, j] <- .stars(pmat[i, j])
  thresh  <- -log10(0.05)
  max_val <- max(logp, na.rm = TRUE)
  col_fn  <- circlize::colorRamp2(
    c(0, thresh, max(thresh + 0.01, max_val)),
    c("white", "orange", "red"))
  ht <- Heatmap(
    logp, name = "-log10(p.adj)",
    col  = col_fn,
    cell_fun = function(j, i, x, y, width, height, fill) {
      s <- stars_mat[i, j]
      if (nchar(s) > 0)
        grid.text(s, x, y, gp = gpar(fontsize = if (is_grouped) 9 else 6))
    },
    cluster_rows = FALSE, cluster_columns = FALSE,
    row_names_gp    = gpar(fontsize = if (is_grouped) 8 else 6),
    column_names_gp = gpar(fontsize = if (is_grouped) 8 else 6),
    column_title    = paste0(nm1, " vs ", nm2, ": significance (", pair_type, " pairs)"),
    heatmap_legend_param = list(
      title_gp  = gpar(fontsize = 8),
      labels_gp = gpar(fontsize = 7))
  )
  draw(ht)
  invisible(NULL)
}

.cc_stats <- function(cc_obj, cc_obj_all = NULL, is_grouped = FALSE) {
  pair_label  <- if (is_grouped) "cell-group" else "cell-type"
  n_levels    <- nlevels(cc_obj@idents)
  max_pairs   <- n_levels * n_levels
  n_lr        <- sum(cc_obj@net$count)
  n_pairs     <- sum(cc_obj@net$count > 0)
  n_pw        <- length(cc_obj@netP$pathways)
  if (!is.null(cc_obj_all)) {
    list(n_lr = n_lr, n_pairs = n_pairs, n_pw = n_pw,
         n_lr_all = sum(cc_obj_all@net$count),
         n_pairs_all = sum(cc_obj_all@net$count > 0),
         n_pw_all = length(cc_obj_all@netP$pathways),
         max_pairs = max_pairs, pair_label = pair_label)
  } else {
    list(n_lr = n_lr, n_pairs = n_pairs, n_pw = n_pw,
         n_lr_all = NULL, n_pairs_all = NULL, n_pw_all = NULL,
         max_pairs = max_pairs, pair_label = pair_label)
  }
}

# For single-condition overview heatmaps (Fig1-4, Fig14-17):
# netVisual_heatmap — matrix of LR count or comm. probability per sender-receiver pair.
interaction_note <- function(cc_obj, nm, cc_obj_all = NULL, is_grouped = FALSE) {
  s <- .cc_stats(cc_obj, cc_obj_all, is_grouped)
  note <- if (sfx == "pfilt" && !is.null(cc_obj_all)) {
    n_filt_lr <- s$n_lr_all - s$n_lr
    n_filt_pw <- s$n_pw_all - s$n_pw
    pct_lr    <- paste0(round(100 * n_filt_lr / max(s$n_lr_all, 1)), "%")
    paste0(
      "*", nm, ":* ",
      "*ligand\u2013receptor (LR) interactions:* ", s$n_lr, "/", s$n_lr_all, " retained (",
        n_filt_lr, " removed, ", pct_lr, "); ",
      "*active ", s$pair_label, " pairs:* ", s$n_pairs, "/", s$n_pairs_all,
        " (", s$n_pairs_all - s$n_pairs, " lost all significant interactions); ",
      "*pathways:* ", s$n_pw, "/", s$n_pw_all,
        " (", n_filt_pw, " lost all significant ligand\u2013receptor pairs). ",
      "*Filter:* p\u00a0\u22640.05, ", FILT_METHOD, "."
    )
  } else {
    paste0(
      "*", nm, ":* ",
      "*ligand\u2013receptor (LR) interactions:* ", s$n_lr, "; ",
      "*active ", s$pair_label, " pairs:* ", s$n_pairs, "/", s$max_pairs, "; ",
      "*pathways:* ", s$n_pw, ". ",
      "*No p-value filtering applied.*"
    )
  }
  cat(paste0(note, "\n\n"))
}

# For signaling role heatmaps (Fig6-7):
# netAnalysis_signalingRole_heatmap — rows: pathways (union of both conditions),
# cols: cell types, values: outgoing/incoming contribution score.
signaling_role_note <- function(pattern) {
  pw1   <- cc[[nm1]]@netP$pathways
  pw2   <- cc[[nm2]]@netP$pathways
  pw_un <- union(pw1, pw2)
  filt_txt <- if (sfx == "pfilt")
    paste0("pathways retained after p\u00a0\u22640.05 filtering (", FILT_METHOD, ")")
  else
    "all pathways, no p-value filtering"
  cat(paste0(
    "*Rows:* union of ", length(pw_un), " pathways (",
    nm1, ": ", length(pw1), "; ", nm2, ": ", length(pw2), "). ",
    "*Columns:* cell types. ",
    "*Values:* ", pattern, " signaling contribution score per cell type per pathway ",
    "(from netAnalysis_computeCentrality on communication probability). ",
    "*Pathway set:* ", filt_txt, ".\n\n"
  ))
  pair_note()
}

# For compareInteractions bar charts (Fig8-9, Fig18-19):
# Shows total LR count or total communication probability per condition (one bar each).
# Statistical test: Wilcoxon rank-sum on per-sample totals.
compare_note <- function(measure = "count", is_grouped = FALSE) {
  obj1 <- if (is_grouped) grouped[[nm1]] else cc[[nm1]]
  obj2 <- if (is_grouped) grouped[[nm2]] else cc[[nm2]]
  obj1_all <- if (sfx == "pfilt") (if (is_grouped) grouped_all[[nm1]] else cc_all[[nm1]]) else NULL
  obj2_all <- if (sfx == "pfilt") (if (is_grouped) grouped_all[[nm2]] else cc_all[[nm2]]) else NULL
  unit <- if (measure == "count") "total ligand\u2013receptor (LR) interactions" else "total communication probability"
  cat(paste0(
    "*Bars show ", unit, " per condition (summed across all ",
    if (is_grouped) "cell-group" else "cell-type", " pairs).*\n\n"
  ))
  .total_test_note(measure)
  interaction_note(obj1, nm1, obj1_all, is_grouped)
  interaction_note(obj2, nm2, obj2_all, is_grouped)
}

# For differential chord/arrow plots (Fig10-11, Fig20-21):
# netVisual_diffInteraction — arrow/edge direction and thickness encode
# the signed difference: positive (red) = more in nm1, negative (blue) = more in nm2.
# Statistical test: Wilcoxon rank-sum per sender-receiver pair with BH correction.
diff_chord_note <- function(measure = "count", is_grouped = FALSE) {
  unit <- if (measure == "count") "ligand\u2013receptor (LR) interaction count" else "communication probability"
  filt_note <- if (sfx == "pfilt")
    " Only ligand\u2013receptor interactions significant by CellChat permutation test (p\u00a0\u22640.05) are included."
  else ""
  cat(paste0(
    "Differential ", unit, " (", nm1, " \u2212 ", nm2, ") per ",
    if (is_grouped) "cell-group" else "cell-type",
    " pair. Red/outgoing arrows: higher in ", nm1,
    "; blue/outgoing arrows: higher in ", nm2,
    ". Edge width scaled by magnitude.", filt_note, "\n\n"
  ))
  if (is_grouped) grouped_pair_note() else pair_note()
}

# For differential heatmaps (Fig12-13, Fig22-23):
# netVisual_heatmap(merged_cc) — matrix of signed difference per sender-receiver pair.
diff_heatmap_note <- function(measure = "count", is_grouped = FALSE) {
  unit <- if (measure == "count") "ligand\u2013receptor (LR) interaction count" else "communication probability"
  filt_note <- if (sfx == "pfilt")
    " Only ligand\u2013receptor interactions significant by CellChat permutation test (p\u00a0\u22640.05) are included."
  else ""
  cat(paste0(
    "Differential ", unit, " per ",
    if (is_grouped) "cell-group" else "cell-type",
    " pair (", nm1, " \u2212 ", nm2,
    "). Positive (red): higher in ", nm1,
    "; negative (blue): higher in ", nm2,
    ".", filt_note, "\n\n"
  ))
  if (is_grouped) grouped_pair_note() else pair_note()
}

pair_note <- function() {
  interaction_note(cc[[nm1]], nm1, if (sfx == "pfilt") cc_all[[nm1]] else NULL,
                   is_grouped = FALSE)
  interaction_note(cc[[nm2]], nm2, if (sfx == "pfilt") cc_all[[nm2]] else NULL,
                   is_grouped = FALSE)
}

grouped_note <- function(nm) {
  interaction_note(grouped[[nm]], nm, if (sfx == "pfilt") grouped_all[[nm]] else NULL,
                   is_grouped = TRUE)
}

grouped_pair_note <- function() {
  grouped_note(nm1)
  grouped_note(nm2)
}
```

```{r load_so}
so <- readRDS(file = paste0(input_var_dir, params$input_rds))
cc          <- so@misc$cc[[sfx]]
cc_all      <- so@misc$cc$all
merged_cc   <- so@misc$cc_merged[[sfx]][[params$pair]]
grouped     <- so@misc$cc_grouped[[sfx]]
grouped_all <- so@misc$cc_grouped$all
grouped_merged <- so@misc$cc_grouped_merged[[sfx]][[params$pair]]
ranknet_obj <- so@misc$cc_ranknet_merged[[params$pair]]
has_ranknet_stats <- !is.null(ranknet_obj)
```

# Single condition overviews

## Overview heatmaps

```{r overview_render, include=FALSE}
overview_items <- list()
for (nm in c(nm1, nm2)) {
  outdir <- file.path(plots_outdir, "overview", nm)
  dir.create(outdir, showWarnings = FALSE, recursive = TRUE)
  for (measure in c("count", "weight")) {
    lbl_m    <- if (measure == "count") "Interaction counts" else "Interaction weight"
    fstem    <- paste0(nm, ".", sfx, ".heatmap_", measure)
    ttl      <- paste0(nm, " (", sfx, "): ", lbl_m)
    plt      <- if (measure == "count") netVisual_heatmap(cc[[nm]]) else
                                        netVisual_heatmap(cc[[nm]], measure = "weight")
    grob     <- grid.grabExpr(draw_grid_plot(plt, ttl))
    overview_items[[length(overview_items) + 1]] <-
      list(nm = nm, measure = measure, lbl = lbl_m, grob = grob,
           suffix = paste0(nm, ".", measure), plot_dir = outdir)
  }
}
```

```{r overview_display, results='asis'}
for (item in overview_items) {
  .fig_counter <- .fig_counter + 1L
  block_prefix <- paste0(.fig_counter, ".", pfx)
  lbl <- paste0(block_prefix, " \u2014 ", item$nm, ": ", item$lbl)
  plot_bundle <- report_plot_bundle(
    item$grob,
    block_prefix = block_prefix,
    suffix = item$suffix,
    width = 7,
    height = 7,
    plot_dir = item$plot_dir,
    data = interaction_matrix_df(cc[[item$nm]], item$nm, item$measure)
  )
  cat(paste0("\n\n### ", lbl, "\n\n"))
  cat(as.character(report_bundle_asset_bar(
    report_figure_name(block_prefix, suffix = item$suffix),
    plot_bundle,
    png_label = "Download PNG",
    pdf_label = "Download PDF",
    tsv_label = "Download TSV"
  )), "\n")
  cat(paste0('<img src="', plot_bundle$png_href, '" width="672" />', "\n\n"))
  interaction_note(cc[[item$nm]], item$nm,
                   if (sfx == "pfilt") cc_all[[item$nm]] else NULL)
}
```

# Comparison: `r nm1` vs `r nm2`

## Pathway ranking (rankNet)

```{r ranknet, fig.width=12, fig.height=7, results='asis'}
pw <- if (sfx == "pfilt") intersect(union(cc[[nm1]]@netP$pathways, cc[[nm2]]@netP$pathways),
                                    merged_cc@netP$pathways) else NULL
n_pw_all   <- length(union(so@misc$cc$all[[nm1]]@netP$pathways,
                            so@misc$cc$all[[nm2]]@netP$pathways))
n_pw_filter <- if (is.null(pw)) n_pw_all else length(pw)
n1 <- length(so@misc[["cellchat_per_sample"]][[nm1]])
n2 <- length(so@misc[["cellchat_per_sample"]][[nm2]])

# rankNet(do.stat=TRUE) requires comparable network dimensions.  Subset both
# condition-level CellChat objects to the intersection of cell types before
# merging them for the pathway ranking test.
.quiet_cellchat <- function(expr) {
  value <- NULL
  invisible(capture.output(value <- eval.parent(substitute(expr))))
  value
}
ranknet_idents1 <- levels(cc[[nm1]]@idents)
ranknet_idents2 <- levels(cc[[nm2]]@idents)
ranknet_common_idents <- intersect(ranknet_idents1, ranknet_idents2)
ranknet_removed_idents <- union(setdiff(ranknet_idents1, ranknet_common_idents),
                                setdiff(ranknet_idents2, ranknet_common_idents))
ranknet_merged <- .quiet_cellchat(
  mergeCellChat(
    setNames(list(
      subsetCellChat(cc[[nm1]], idents.use = ranknet_common_idents),
      subsetCellChat(cc[[nm2]], idents.use = ranknet_common_idents)
    ), c(nm1, nm2)),
    add.names = c(nm1, nm2)
  )
)

ranknet_rel <- rankNet(ranknet_merged, mode = "comparison", measure = "weight",
                       stacked = TRUE, do.stat = TRUE, paired.test = TRUE,
                       signaling = pw, thresh = thresh, return.data = TRUE)
ranknet_abs <- rankNet(ranknet_merged, mode = "comparison", measure = "weight",
                       stacked = FALSE, do.stat = TRUE, paired.test = TRUE,
                       signaling = pw, thresh = thresh, return.data = TRUE)
gg1 <- ranknet_rel$gg.obj
gg2 <- ranknet_abs$gg.obj

# CellChat's rankNet removes rows for pathways thresholded to zero but retains
# them in the factor levels (R does not auto-drop unused factor levels when
# removing rows).  This produces empty y-axis entries with no bars, most
# conspicuous in the relative (fill) panel.  Drop them here so only pathways
# with at least one non-zero contribution appear in the plot.
.drop_empty_pathway_levels <- function(gg) {
  d <- gg$data
  if (is.null(d) || !"name" %in% names(d) || !"contribution" %in% names(d))
    return(gg)
  active <- with(d, names(which(tapply(contribution, name,
                                       function(x) any(!is.na(x) & x > 0)))))
  if (length(active) == nlevels(d$name)) return(gg)
  gg$data <- droplevels(d[as.character(d$name) %in% active, , drop = FALSE])
  gg
}

# rankNet's relative panel uses position="fill" on tiny raw probabilities.
# For very small non-zero pathways (for example DHEA), the internal relative
# values can be rounded to zero, yielding no visible bar.  Compute the relative
# heights explicitly from the raw contribution so non-zero pathways are drawn.
.use_explicit_relative_ranknet <- function(gg) {
  d <- gg$data
  if (is.null(d) || !"name" %in% names(d) || !"contribution" %in% names(d))
    return(gg)
  total <- ave(d$contribution, d$name, FUN = sum)
  d$contribution.relative.plot <- ifelse(total > 0, d$contribution / total, 0)
  gg$data <- d
  gg$mapping <- ggplot2::aes(x = name, y = contribution.relative.plot, fill = group)
  gg$layers[[1]]$position <- ggplot2::position_stack()
  gg$labels$y <- "Relative information flow"
  gg
}
gg1 <- .use_explicit_relative_ranknet(gg1)
gg1 <- .drop_empty_pathway_levels(gg1)
gg2 <- .drop_empty_pathway_levels(gg2)
# Actual number of pathways with data (may be less than n_pw_filter if some
# were thresholded to zero in the merged object).
n_pw_shown <- if (!is.null(gg1$data) && "name" %in% names(gg1$data))
                nlevels(gg1$data$name) else n_pw_filter
n_pw_ghost  <- n_pw_filter - n_pw_shown  # in pw filter but no data in merged obj

# CellChat rankNet p-values across matched sender-receiver cell-type pairs.
pw_stats <- .ranknet_stats(ranknet_rel$signaling.contribution, thresh = 0.05)
if (!is.null(pw_stats)) {
  gg1 <- .add_ranknet_pvalues(gg1, pw_stats)
  gg2 <- .add_ranknet_pvalues(gg2, pw_stats)
  gg1 <- .color_ranknet_labels(gg1, pw_stats, thresh = 0.05)
  gg2 <- .color_ranknet_labels(gg2, pw_stats, thresh = 0.05)
}

ranknet_pdf_file <- paste0(params$pair, ".", sfx, ".ranknet.pdf")
ranknet_pdf_path <- file.path(plots_outdir, "comparisons", ranknet_pdf_file)
ranknet_png_file <- paste0(params$pair, ".", sfx, ".ranknet.png")
ranknet_png_path <- file.path(plots_outdir, "comparisons", ranknet_png_file)
ranknet_tsv_file <- paste0(params$pair, ".", sfx, ".ranknet.tsv")
ranknet_tsv_path <- file.path(plots_outdir, "comparisons", ranknet_tsv_file)
dir.create(file.path(plots_outdir, "comparisons"), showWarnings = FALSE, recursive = TRUE)
combined_gg <- (gg1 + gg2) + plot_annotation(title = paste0(params$pair, ": Pathway ranking (", sfx, ")"))
ggsave(ranknet_pdf_path, combined_gg, width = 12, height = 7)
ggsave(ranknet_png_path, combined_gg, width = 12, height = 7)
ranknet_rel_df <- gg_data_df(gg1, "relative_information_flow")
ranknet_abs_df <- gg_data_df(gg2, "absolute_information_flow")
ranknet_cols <- union(names(ranknet_rel_df), names(ranknet_abs_df))
ranknet_rel_df[setdiff(ranknet_cols, names(ranknet_rel_df))] <- NA
ranknet_abs_df[setdiff(ranknet_cols, names(ranknet_abs_df))] <- NA
write_tsv(rbind(ranknet_rel_df[ranknet_cols], ranknet_abs_df[ranknet_cols]), ranknet_tsv_path)
fig_label(paste0(params$pair, ": Pathway ranking"),
          png_href = report_href(file.path("comparisons", ranknet_png_file)),
          png_fname = ranknet_png_file,
          pdf_href = report_href(file.path("comparisons", ranknet_pdf_file)),
          pdf_fname = ranknet_pdf_file,
          tsv_href = report_href(file.path("comparisons", ranknet_tsv_file)),
          tsv_fname = ranknet_tsv_file)
print(combined_gg)

# --- Note ---
pw_note <- if (sfx == "pfilt") {
  ghost_note <- if (n_pw_ghost > 0)
    paste0("; ", n_pw_ghost,
           " additional pathway(s) passed the per-condition filter but had",
           " no active interactions in the merged object (removed from plot)")
  else ""
  paste0("Pathways shown: ", n_pw_shown,
         " with at least one significant ligand\u2013receptor interaction ",
         "(CellChat permutation test, p\u00a0\u22640.05; ",
         n_pw_all - n_pw_filter, " pathways removed as entirely non-significant",
         ghost_note, ")")
} else {
  paste0("Pathways shown: ", n_pw_shown, " (all pathways, unfiltered)")
}
panel_note <- paste0(
  "Left panel: relative information flow per pathway and condition. ",
  "Right panel: absolute information flow using a \u2212\u2009(1/log) ",
  "transformation of raw probability for visual scaling. ",
  "Because the two panels use different scales, bars that look similar in the ",
  "absolute panel may appear very unequal in the relative panel, and vice versa.")
intersection_note <- if (length(ranknet_removed_idents) > 0) {
  paste0("For rankNet statistics, ", length(ranknet_removed_idents),
         " cell type(s) were removed before merging because they were missing ",
         "from one condition network: ",
         paste(ranknet_removed_idents, collapse = ", "), ".")
} else {
  ""
}
if (!is.null(pw_stats)) {
  stat_note <- paste0(
    "Between-condition test: CellChat rankNet(do.stat=TRUE) Wilcoxon signed-rank test ",
    "for condition-level network comparison of matched sender-receiver cell-type-pair ",
    "communication probabilities after subsetting both condition networks to their ",
    "shared cell types (", length(ranknet_common_idents), " cell types). ",
    "Raw p-values were adjusted across tested pathways using Benjamini-Hochberg FDR; ",
    "FDR\u00a0<\u00a00.05 is used as the pathway-label threshold. ",
    pw_stats$n_tested, " pathway(s) were tested; ",
    pw_stats$n_sig, " pathway(s) pass this threshold. ",
    "Pathway label colors: **red** = stronger in ", nm1, " (rankNet FDR\u00a0<\u00a00.05)",
    "; **blue** = stronger in ", nm2, " (rankNet FDR\u00a0<\u00a00.05)",
    "; black = not significant at this threshold.")
} else {
  stat_note <- "Between-condition test: CellChat rankNet(do.stat=TRUE) could not be computed. Pathway label colors are not significance-coded."
}
cat(paste0("*", pw_note, ".*\n\n"))
cat(paste0("*", panel_note, "*\n\n"))
if (nchar(intersection_note) > 0)
  cat(paste0("*", intersection_note, "*\n\n"))
cat(paste0("*", stat_note, "*\n\n"))
```

## Signaling role heatmaps

```{r signaling_role_heatmaps, fig.width=12, fig.height=9, results='asis'}
sig_out_pdf_file <- paste0(params$pair, ".", sfx, ".sig_role_outgoing.pdf")
sig_out_pdf_path <- file.path(plots_outdir, "comparisons", sig_out_pdf_file)
sig_out_png_file <- paste0(params$pair, ".", sfx, ".sig_role_outgoing.png")
sig_out_png_path <- file.path(plots_outdir, "comparisons", sig_out_png_file)
sig_out_tsv_file <- paste0(params$pair, ".", sfx, ".sig_role_outgoing.tsv")
sig_out_tsv_path <- file.path(plots_outdir, "comparisons", sig_out_tsv_file)
write_tsv(rbind(signaling_role_df(cc[[nm1]], nm1, "outgoing"),
                signaling_role_df(cc[[nm2]], nm2, "outgoing")), sig_out_tsv_path)
fig_label(paste0(params$pair, ": Signaling role heatmap (outgoing)"),
          png_href = report_href(file.path("comparisons", sig_out_png_file)),
          png_fname = sig_out_png_file,
          pdf_href = report_href(file.path("comparisons", sig_out_pdf_file)),
          pdf_fname = sig_out_pdf_file,
          tsv_href = report_href(file.path("comparisons", sig_out_tsv_file)),
          tsv_fname = sig_out_tsv_file)
make_pair_ht(cc[[nm1]], cc[[nm2]], left_title = nm1, right_title = nm2,
             pattern = "outgoing", width = 10, height = 12, gap_cm = 1.5,
             pdf_path = sig_out_pdf_path, png_path = sig_out_png_path)
signaling_role_note("Outgoing")
sig_in_pdf_file <- paste0(params$pair, ".", sfx, ".sig_role_incoming.pdf")
sig_in_pdf_path <- file.path(plots_outdir, "comparisons", sig_in_pdf_file)
sig_in_png_file <- paste0(params$pair, ".", sfx, ".sig_role_incoming.png")
sig_in_png_path <- file.path(plots_outdir, "comparisons", sig_in_png_file)
sig_in_tsv_file <- paste0(params$pair, ".", sfx, ".sig_role_incoming.tsv")
sig_in_tsv_path <- file.path(plots_outdir, "comparisons", sig_in_tsv_file)
write_tsv(rbind(signaling_role_df(cc[[nm1]], nm1, "incoming"),
                signaling_role_df(cc[[nm2]], nm2, "incoming")), sig_in_tsv_path)
fig_label(paste0(params$pair, ": Signaling role heatmap (incoming)"),
          png_href = report_href(file.path("comparisons", sig_in_png_file)),
          png_fname = sig_in_png_file,
          pdf_href = report_href(file.path("comparisons", sig_in_pdf_file)),
          pdf_fname = sig_in_pdf_file,
          tsv_href = report_href(file.path("comparisons", sig_in_tsv_file)),
          tsv_fname = sig_in_tsv_file)
make_pair_ht(cc[[nm1]], cc[[nm2]], left_title = nm1, right_title = nm2,
             pattern = "incoming", width = 10, height = 12, gap_cm = 1.5,
             pdf_path = sig_in_pdf_path, png_path = sig_in_png_path)
signaling_role_note("Incoming")
```

## Comparison plots

```{r comparison_plots, fig.width=12, fig.height=10, results='asis'}
outdir <- file.path(plots_outdir, "comparisons")
dir.create(outdir, showWarnings = FALSE, recursive = TRUE)

cmp_pdf <- function(fname) report_href(file.path("comparisons", fname))

ci_count_pdf <- paste0(params$pair, ".", sfx, ".compareInteractions_count.pdf")
ci_count_png <- paste0(params$pair, ".", sfx, ".compareInteractions_count.png")
ci_count_tsv <- paste0(params$pair, ".", sfx, ".compareInteractions_count.tsv")
fig_label(paste0(params$pair, ": Number of interactions"),
          png_href = cmp_pdf(ci_count_png), png_fname = ci_count_png,
          pdf_href = cmp_pdf(ci_count_pdf), pdf_fname = ci_count_pdf,
          tsv_href = cmp_pdf(ci_count_tsv), tsv_fname = ci_count_tsv)
gg1 <- compareInteractions(merged_cc, show.legend = FALSE, group = c(1,2)) +
  ggtitle(paste0(params$pair, " (", sfx, "): Number of interactions"))
ci_colors <- .extract_ci_colors(gg1)
print(gg1)
ggsave(file.path(outdir, ci_count_pdf), plot = gg1, width = 7, height = 5)
ggsave(file.path(outdir, ci_count_png), plot = gg1, width = 7, height = 5)
write_tsv(rbind(interaction_matrix_df(cc[[nm1]], nm1, "count"),
                interaction_matrix_df(cc[[nm2]], nm2, "count")),
          file.path(outdir, ci_count_tsv))
compare_note(measure = "count", is_grouped = FALSE)
ci_count_sample_pdf <- paste0(params$pair, ".", sfx, ".compareInteractions_count_per_sample.pdf")
ci_count_sample_png <- paste0(params$pair, ".", sfx, ".compareInteractions_count_per_sample.png")
ci_count_sample_tsv <- paste0(params$pair, ".", sfx, ".compareInteractions_count_per_sample.tsv")
fig_label(paste0(params$pair, ": Number of interactions per sample"),
          png_href = cmp_pdf(ci_count_sample_png), png_fname = ci_count_sample_png,
          pdf_href = cmp_pdf(ci_count_sample_pdf), pdf_fname = ci_count_sample_pdf,
          tsv_href = cmp_pdf(ci_count_sample_tsv), tsv_fname = ci_count_sample_tsv)
ps_gg <- .persample_barplot(measure = "count", is_grouped = FALSE, colors = ci_colors,
                            tsv_path = file.path(outdir, ci_count_sample_tsv))
if (!is.null(ps_gg)) {
  ggsave(file.path(outdir, ci_count_sample_pdf), plot = ps_gg, width = 7, height = 5)
  ggsave(file.path(outdir, ci_count_sample_png), plot = ps_gg, width = 7, height = 5)
}

ci_weight_pdf <- paste0(params$pair, ".", sfx, ".compareInteractions_weight.pdf")
ci_weight_png <- paste0(params$pair, ".", sfx, ".compareInteractions_weight.png")
ci_weight_tsv <- paste0(params$pair, ".", sfx, ".compareInteractions_weight.tsv")
fig_label(paste0(params$pair, ": Interaction strength"),
          png_href = cmp_pdf(ci_weight_png), png_fname = ci_weight_png,
          pdf_href = cmp_pdf(ci_weight_pdf), pdf_fname = ci_weight_pdf,
          tsv_href = cmp_pdf(ci_weight_tsv), tsv_fname = ci_weight_tsv)
gg2 <- compareInteractions(merged_cc, show.legend = FALSE, group = c(1,2), measure = "weight") +
  ggtitle(paste0(params$pair, " (", sfx, "): Interaction strength"))
print(gg2)
ggsave(file.path(outdir, ci_weight_pdf), plot = gg2, width = 7, height = 5)
ggsave(file.path(outdir, ci_weight_png), plot = gg2, width = 7, height = 5)
write_tsv(rbind(interaction_matrix_df(cc[[nm1]], nm1, "weight"),
                interaction_matrix_df(cc[[nm2]], nm2, "weight")),
          file.path(outdir, ci_weight_tsv))
compare_note(measure = "weight", is_grouped = FALSE)
ci_weight_sample_pdf <- paste0(params$pair, ".", sfx, ".compareInteractions_weight_per_sample.pdf")
ci_weight_sample_png <- paste0(params$pair, ".", sfx, ".compareInteractions_weight_per_sample.png")
ci_weight_sample_tsv <- paste0(params$pair, ".", sfx, ".compareInteractions_weight_per_sample.tsv")
fig_label(paste0(params$pair, ": Interaction strength per sample"),
          png_href = cmp_pdf(ci_weight_sample_png), png_fname = ci_weight_sample_png,
          pdf_href = cmp_pdf(ci_weight_sample_pdf), pdf_fname = ci_weight_sample_pdf,
          tsv_href = cmp_pdf(ci_weight_sample_tsv), tsv_fname = ci_weight_sample_tsv)
ps_gg <- .persample_barplot(measure = "weight", is_grouped = FALSE, colors = ci_colors,
                            tsv_path = file.path(outdir, ci_weight_sample_tsv))
if (!is.null(ps_gg)) {
  ggsave(file.path(outdir, ci_weight_sample_pdf), plot = ps_gg, width = 7, height = 5)
  ggsave(file.path(outdir, ci_weight_sample_png), plot = ps_gg, width = 7, height = 5)
}

di_count_pdf <- paste0(params$pair, ".", sfx, ".diffInteraction_count.pdf")
di_count_png <- paste0(params$pair, ".", sfx, ".diffInteraction_count.png")
di_count_tsv <- paste0(params$pair, ".", sfx, ".diffInteraction_count.tsv")
fig_label(paste0(params$pair, ": Differential number of interactions"),
          png_href = cmp_pdf(di_count_png), png_fname = di_count_png,
          pdf_href = cmp_pdf(di_count_pdf), pdf_fname = di_count_pdf,
          tsv_href = cmp_pdf(di_count_tsv), tsv_fname = di_count_tsv)
write_tsv(merged_diff_df(merged_cc, "count"), file.path(outdir, di_count_tsv))
pdf(file.path(outdir, di_count_pdf), width = 12, height = 10)
op <- par(mar = c(2, 2, 6, 2), xpd = NA)
netVisual_diffInteraction(merged_cc, weight.scale = TRUE)
par(op); invisible(dev.off())
png(file.path(outdir, di_count_png), width = 1200, height = 1000)
op <- par(mar = c(2, 2, 6, 2), xpd = NA)
netVisual_diffInteraction(merged_cc, weight.scale = TRUE)
try(title(paste0(params$pair, " (", sfx, "): Differential number of interactions"),
      cex.main = 1.4, line = 2), silent = TRUE)
par(op); invisible(dev.off())
op <- par(mar = c(2, 2, 6, 2), xpd = NA)
netVisual_diffInteraction(merged_cc, weight.scale = TRUE)
try(title(paste0(params$pair, " (", sfx, "): Differential number of interactions"),
      cex.main = 1.4, line = 2), silent = TRUE)
par(op)
diff_chord_note(measure = "count", is_grouped = FALSE)

di_weight_pdf <- paste0(params$pair, ".", sfx, ".diffInteraction_weight.pdf")
di_weight_png <- paste0(params$pair, ".", sfx, ".diffInteraction_weight.png")
di_weight_tsv <- paste0(params$pair, ".", sfx, ".diffInteraction_weight.tsv")
fig_label(paste0(params$pair, ": Differential interaction strengths"),
          png_href = cmp_pdf(di_weight_png), png_fname = di_weight_png,
          pdf_href = cmp_pdf(di_weight_pdf), pdf_fname = di_weight_pdf,
          tsv_href = cmp_pdf(di_weight_tsv), tsv_fname = di_weight_tsv)
write_tsv(merged_diff_df(merged_cc, "weight"), file.path(outdir, di_weight_tsv))
pdf(file.path(outdir, di_weight_pdf), width = 12, height = 10)
op <- par(mar = c(2, 2, 6, 2), xpd = NA)
netVisual_diffInteraction(merged_cc, weight.scale = TRUE, measure = "weight")
par(op); invisible(dev.off())
png(file.path(outdir, di_weight_png), width = 1200, height = 1000)
op <- par(mar = c(2, 2, 6, 2), xpd = NA)
netVisual_diffInteraction(merged_cc, weight.scale = TRUE, measure = "weight")
try(title(paste0(params$pair, " (", sfx, "): Differential interaction strengths"),
      cex.main = 1.4, line = 2), silent = TRUE)
par(op); invisible(dev.off())
op <- par(mar = c(2, 2, 6, 2), xpd = NA)
netVisual_diffInteraction(merged_cc, weight.scale = TRUE, measure = "weight")
try(title(paste0(params$pair, " (", sfx, "): Differential interaction strengths"),
      cex.main = 1.4, line = 2), silent = TRUE)
par(op)
diff_chord_note(measure = "weight", is_grouped = FALSE)

dh_count_pdf <- paste0(params$pair, ".", sfx, ".diff_heatmap_count.pdf")
dh_count_png <- paste0(params$pair, ".", sfx, ".diff_heatmap_count.png")
dh_count_tsv <- paste0(params$pair, ".", sfx, ".diff_heatmap_count.tsv")
fig_label(paste0(params$pair, ": Differential interaction counts heatmap"),
          png_href = cmp_pdf(dh_count_png), png_fname = dh_count_png,
          pdf_href = cmp_pdf(dh_count_pdf), pdf_fname = dh_count_pdf,
          tsv_href = cmp_pdf(dh_count_tsv), tsv_fname = dh_count_tsv)
write_tsv(merged_diff_df(merged_cc, "count"), file.path(outdir, dh_count_tsv))
plt   <- netVisual_heatmap(merged_cc)
ttl   <- paste0(params$pair, " (", sfx, "): Differential interaction counts")
draw_grid_plot(plt, ttl)
pdf(file.path(outdir, dh_count_pdf), width = 7, height = 7)
draw_grid_plot(plt, ttl); invisible(dev.off())
png(file.path(outdir, dh_count_png), width = 700, height = 700)
draw_grid_plot(plt, ttl); invisible(dev.off())
diff_heatmap_note(measure = "count", is_grouped = FALSE)

dh_weight_pdf <- paste0(params$pair, ".", sfx, ".diff_heatmap_weight.pdf")
dh_weight_png <- paste0(params$pair, ".", sfx, ".diff_heatmap_weight.png")
dh_weight_tsv <- paste0(params$pair, ".", sfx, ".diff_heatmap_weight.tsv")
fig_label(paste0(params$pair, ": Differential interaction weights heatmap"),
          png_href = cmp_pdf(dh_weight_png), png_fname = dh_weight_png,
          pdf_href = cmp_pdf(dh_weight_pdf), pdf_fname = dh_weight_pdf,
          tsv_href = cmp_pdf(dh_weight_tsv), tsv_fname = dh_weight_tsv)
write_tsv(merged_diff_df(merged_cc, "weight"), file.path(outdir, dh_weight_tsv))
plt   <- netVisual_heatmap(merged_cc, measure = "weight")
ttl   <- paste0(params$pair, " (", sfx, "): Differential interaction weights")
draw_grid_plot(plt, ttl)
pdf(file.path(outdir, dh_weight_pdf), width = 7, height = 7)
draw_grid_plot(plt, ttl); invisible(dev.off())
png(file.path(outdir, dh_weight_png), width = 700, height = 700)
draw_grid_plot(plt, ttl); invisible(dev.off())
diff_heatmap_note(measure = "weight", is_grouped = FALSE)
```

# Grouped cell types

## Grouped overview heatmaps

```{r grouped_overview_render, include=FALSE}
grouped_overview_items <- list()
for (nm in c(nm1, nm2)) {
  outdir <- file.path(plots_outdir, "overview_grouped", nm)
  dir.create(outdir, showWarnings = FALSE, recursive = TRUE)
  for (measure in c("count", "weight")) {
    lbl_m    <- if (measure == "count") "Interaction counts" else "Interaction weight"
    fstem    <- paste0(nm, ".", sfx, ".grouped_heatmap_", measure)
    ttl      <- paste0(nm, " (", sfx, "): ", lbl_m, " (grouped)")
    plt      <- if (measure == "count") netVisual_heatmap(grouped[[nm]]) else
                                        netVisual_heatmap(grouped[[nm]], measure = "weight")
    grob     <- grid.grabExpr(draw_grid_plot(plt, ttl))
    grouped_overview_items[[length(grouped_overview_items) + 1]] <-
      list(nm = nm, measure = measure, lbl = lbl_m, grob = grob,
           suffix = paste0(nm, ".grouped.", measure), plot_dir = outdir)
  }
}
```

```{r grouped_overview_display, results='asis'}
for (item in grouped_overview_items) {
  .fig_counter <- .fig_counter + 1L
  block_prefix <- paste0(.fig_counter, ".", pfx)
  lbl <- paste0(block_prefix, " \u2014 ", item$nm, " grouped: ", item$lbl)
  plot_bundle <- report_plot_bundle(
    item$grob,
    block_prefix = block_prefix,
    suffix = item$suffix,
    width = 7,
    height = 7,
    plot_dir = item$plot_dir,
    data = interaction_matrix_df(grouped[[item$nm]], item$nm, item$measure)
  )
  cat(paste0("\n\n### ", lbl, "\n\n"))
  cat(as.character(report_bundle_asset_bar(
    report_figure_name(block_prefix, suffix = item$suffix),
    plot_bundle,
    png_label = "Download PNG",
    pdf_label = "Download PDF",
    tsv_label = "Download TSV"
  )), "\n")
  cat(paste0('<img src="', plot_bundle$png_href, '" width="672" />', "\n\n"))
  grouped_note(item$nm)
}
```

## Grouped comparison

```{r grouped_comparison_plots, fig.width=12, fig.height=10, results='asis'}
outdir <- file.path(plots_outdir, "comparisons_grouped")
dir.create(outdir, showWarnings = FALSE, recursive = TRUE)

gcmp_pdf <- function(fname) report_href(file.path("comparisons_grouped", fname))

gci_count_pdf <- paste0(params$pair, ".", sfx, ".grouped_compareInteractions_count.pdf")
gci_count_png <- paste0(params$pair, ".", sfx, ".grouped_compareInteractions_count.png")
gci_count_tsv <- paste0(params$pair, ".", sfx, ".grouped_compareInteractions_count.tsv")
fig_label(paste0(params$pair, " grouped: Number of interactions"),
          png_href = gcmp_pdf(gci_count_png), png_fname = gci_count_png,
          pdf_href = gcmp_pdf(gci_count_pdf), pdf_fname = gci_count_pdf,
          tsv_href = gcmp_pdf(gci_count_tsv), tsv_fname = gci_count_tsv)
gg1 <- compareInteractions(grouped_merged, show.legend = FALSE, group = c(1,2)) +
  ggtitle(paste0(params$pair, " (", sfx, "): Number of interactions (grouped)"))
gci_colors <- .extract_ci_colors(gg1)
print(gg1)
ggsave(file.path(outdir, gci_count_pdf), plot = gg1, width = 7, height = 5)
ggsave(file.path(outdir, gci_count_png), plot = gg1, width = 7, height = 5)
write_tsv(rbind(interaction_matrix_df(grouped[[nm1]], nm1, "count"),
                interaction_matrix_df(grouped[[nm2]], nm2, "count")),
          file.path(outdir, gci_count_tsv))
compare_note(measure = "count", is_grouped = TRUE)
gci_count_sample_pdf <- paste0(params$pair, ".", sfx, ".grouped_compareInteractions_count_per_sample.pdf")
gci_count_sample_png <- paste0(params$pair, ".", sfx, ".grouped_compareInteractions_count_per_sample.png")
gci_count_sample_tsv <- paste0(params$pair, ".", sfx, ".grouped_compareInteractions_count_per_sample.tsv")
fig_label(paste0(params$pair, " grouped: Number of interactions per sample"),
          png_href = gcmp_pdf(gci_count_sample_png), png_fname = gci_count_sample_png,
          pdf_href = gcmp_pdf(gci_count_sample_pdf), pdf_fname = gci_count_sample_pdf,
          tsv_href = gcmp_pdf(gci_count_sample_tsv), tsv_fname = gci_count_sample_tsv)
ps_gg <- .persample_barplot(measure = "count", is_grouped = TRUE, colors = gci_colors,
                            tsv_path = file.path(outdir, gci_count_sample_tsv))
if (!is.null(ps_gg)) {
  ggsave(file.path(outdir, gci_count_sample_pdf), plot = ps_gg, width = 7, height = 5)
  ggsave(file.path(outdir, gci_count_sample_png), plot = ps_gg, width = 7, height = 5)
}

gci_weight_pdf <- paste0(params$pair, ".", sfx, ".grouped_compareInteractions_weight.pdf")
gci_weight_png <- paste0(params$pair, ".", sfx, ".grouped_compareInteractions_weight.png")
gci_weight_tsv <- paste0(params$pair, ".", sfx, ".grouped_compareInteractions_weight.tsv")
fig_label(paste0(params$pair, " grouped: Interaction strength"),
          png_href = gcmp_pdf(gci_weight_png), png_fname = gci_weight_png,
          pdf_href = gcmp_pdf(gci_weight_pdf), pdf_fname = gci_weight_pdf,
          tsv_href = gcmp_pdf(gci_weight_tsv), tsv_fname = gci_weight_tsv)
gg2 <- compareInteractions(grouped_merged, show.legend = FALSE, group = c(1,2), measure = "weight") +
  ggtitle(paste0(params$pair, " (", sfx, "): Interaction strength (grouped)"))
print(gg2)
ggsave(file.path(outdir, gci_weight_pdf), plot = gg2, width = 7, height = 5)
ggsave(file.path(outdir, gci_weight_png), plot = gg2, width = 7, height = 5)
write_tsv(rbind(interaction_matrix_df(grouped[[nm1]], nm1, "weight"),
                interaction_matrix_df(grouped[[nm2]], nm2, "weight")),
          file.path(outdir, gci_weight_tsv))
compare_note(measure = "weight", is_grouped = TRUE)
gci_weight_sample_pdf <- paste0(params$pair, ".", sfx, ".grouped_compareInteractions_weight_per_sample.pdf")
gci_weight_sample_png <- paste0(params$pair, ".", sfx, ".grouped_compareInteractions_weight_per_sample.png")
gci_weight_sample_tsv <- paste0(params$pair, ".", sfx, ".grouped_compareInteractions_weight_per_sample.tsv")
fig_label(paste0(params$pair, " grouped: Interaction strength per sample"),
          png_href = gcmp_pdf(gci_weight_sample_png), png_fname = gci_weight_sample_png,
          pdf_href = gcmp_pdf(gci_weight_sample_pdf), pdf_fname = gci_weight_sample_pdf,
          tsv_href = gcmp_pdf(gci_weight_sample_tsv), tsv_fname = gci_weight_sample_tsv)
ps_gg <- .persample_barplot(measure = "weight", is_grouped = TRUE, colors = gci_colors,
                            tsv_path = file.path(outdir, gci_weight_sample_tsv))
if (!is.null(ps_gg)) {
  ggsave(file.path(outdir, gci_weight_sample_pdf), plot = ps_gg, width = 7, height = 5)
  ggsave(file.path(outdir, gci_weight_sample_png), plot = ps_gg, width = 7, height = 5)
}

gdi_count_pdf <- paste0(params$pair, ".", sfx, ".grouped_diffInteraction_count.pdf")
gdi_count_png <- paste0(params$pair, ".", sfx, ".grouped_diffInteraction_count.png")
gdi_count_tsv <- paste0(params$pair, ".", sfx, ".grouped_diffInteraction_count.tsv")
fig_label(paste0(params$pair, " grouped: Differential number of interactions"),
          png_href = gcmp_pdf(gdi_count_png), png_fname = gdi_count_png,
          pdf_href = gcmp_pdf(gdi_count_pdf), pdf_fname = gdi_count_pdf,
          tsv_href = gcmp_pdf(gdi_count_tsv), tsv_fname = gdi_count_tsv)
write_tsv(merged_diff_df(grouped_merged, "count"), file.path(outdir, gdi_count_tsv))
pdf(file.path(outdir, gdi_count_pdf), width = 12, height = 10)
op <- par(mar = c(2, 2, 6, 2), xpd = NA)
netVisual_diffInteraction(grouped_merged, weight.scale = TRUE)
par(op); invisible(dev.off())
png(file.path(outdir, gdi_count_png), width = 1200, height = 1000)
op <- par(mar = c(2, 2, 6, 2), xpd = NA)
netVisual_diffInteraction(grouped_merged, weight.scale = TRUE)
try(title(paste0(params$pair, " (", sfx, "): Differential interactions (grouped)"),
      cex.main = 1.4, line = 2), silent = TRUE)
par(op); invisible(dev.off())
op <- par(mar = c(2, 2, 6, 2), xpd = NA)
netVisual_diffInteraction(grouped_merged, weight.scale = TRUE)
try(title(paste0(params$pair, " (", sfx, "): Differential interactions (grouped)"),
      cex.main = 1.4, line = 2), silent = TRUE)
par(op)
diff_chord_note(measure = "count", is_grouped = TRUE)

gdi_weight_pdf <- paste0(params$pair, ".", sfx, ".grouped_diffInteraction_weight.pdf")
gdi_weight_png <- paste0(params$pair, ".", sfx, ".grouped_diffInteraction_weight.png")
gdi_weight_tsv <- paste0(params$pair, ".", sfx, ".grouped_diffInteraction_weight.tsv")
fig_label(paste0(params$pair, " grouped: Differential interaction strengths"),
          png_href = gcmp_pdf(gdi_weight_png), png_fname = gdi_weight_png,
          pdf_href = gcmp_pdf(gdi_weight_pdf), pdf_fname = gdi_weight_pdf,
          tsv_href = gcmp_pdf(gdi_weight_tsv), tsv_fname = gdi_weight_tsv)
write_tsv(merged_diff_df(grouped_merged, "weight"), file.path(outdir, gdi_weight_tsv))
pdf(file.path(outdir, gdi_weight_pdf), width = 12, height = 10)
op <- par(mar = c(2, 2, 6, 2), xpd = NA)
netVisual_diffInteraction(grouped_merged, weight.scale = TRUE, measure = "weight")
par(op); invisible(dev.off())
png(file.path(outdir, gdi_weight_png), width = 1200, height = 1000)
op <- par(mar = c(2, 2, 6, 2), xpd = NA)
netVisual_diffInteraction(grouped_merged, weight.scale = TRUE, measure = "weight")
try(title(paste0(params$pair, " (", sfx, "): Differential strengths (grouped)"),
      cex.main = 1.4, line = 2), silent = TRUE)
par(op); invisible(dev.off())
op <- par(mar = c(2, 2, 6, 2), xpd = NA)
netVisual_diffInteraction(grouped_merged, weight.scale = TRUE, measure = "weight")
try(title(paste0(params$pair, " (", sfx, "): Differential strengths (grouped)"),
      cex.main = 1.4, line = 2), silent = TRUE)
par(op)
diff_chord_note(measure = "weight", is_grouped = TRUE)

gdh_count_pdf <- paste0(params$pair, ".", sfx, ".grouped_diff_heatmap_count.pdf")
gdh_count_png <- paste0(params$pair, ".", sfx, ".grouped_diff_heatmap_count.png")
gdh_count_tsv <- paste0(params$pair, ".", sfx, ".grouped_diff_heatmap_count.tsv")
fig_label(paste0(params$pair, " grouped: Diff interaction counts heatmap"),
          png_href = gcmp_pdf(gdh_count_png), png_fname = gdh_count_png,
          pdf_href = gcmp_pdf(gdh_count_pdf), pdf_fname = gdh_count_pdf,
          tsv_href = gcmp_pdf(gdh_count_tsv), tsv_fname = gdh_count_tsv)
write_tsv(merged_diff_df(grouped_merged, "count"), file.path(outdir, gdh_count_tsv))
plt <- netVisual_heatmap(grouped_merged)
ttl <- paste0(params$pair, " (", sfx, "): Differential interaction counts (grouped)")
draw_grid_plot(plt, ttl)
pdf(file.path(outdir, gdh_count_pdf), width = 7, height = 7)
draw_grid_plot(plt, ttl); invisible(dev.off())
png(file.path(outdir, gdh_count_png), width = 700, height = 700)
draw_grid_plot(plt, ttl); invisible(dev.off())
diff_heatmap_note(measure = "count", is_grouped = TRUE)

gdh_weight_pdf <- paste0(params$pair, ".", sfx, ".grouped_diff_heatmap_weight.pdf")
gdh_weight_png <- paste0(params$pair, ".", sfx, ".grouped_diff_heatmap_weight.png")
gdh_weight_tsv <- paste0(params$pair, ".", sfx, ".grouped_diff_heatmap_weight.tsv")
fig_label(paste0(params$pair, " grouped: Diff interaction weights heatmap"),
          png_href = gcmp_pdf(gdh_weight_png), png_fname = gdh_weight_png,
          pdf_href = gcmp_pdf(gdh_weight_pdf), pdf_fname = gdh_weight_pdf,
          tsv_href = gcmp_pdf(gdh_weight_tsv), tsv_fname = gdh_weight_tsv)
write_tsv(merged_diff_df(grouped_merged, "weight"), file.path(outdir, gdh_weight_tsv))
plt <- netVisual_heatmap(grouped_merged, measure = "weight")
ttl <- paste0(params$pair, " (", sfx, "): Differential interaction weights (grouped)")
draw_grid_plot(plt, ttl)
pdf(file.path(outdir, gdh_weight_pdf), width = 7, height = 7)
draw_grid_plot(plt, ttl); invisible(dev.off())
png(file.path(outdir, gdh_weight_png), width = 700, height = 700)
draw_grid_plot(plt, ttl); invisible(dev.off())
diff_heatmap_note(measure = "weight", is_grouped = TRUE)
```

# Session info

```{r session_info, echo=FALSE}
sessionInfo()
```
