---
title: "`r paste(c(params$report_title_prefix, 'CellChat', if (params$pair %in% names(params$pair_labels)) params$pair_labels[[params$pair]] else params$pair, if (params$sfx %in% names(params$filter_labels)) params$filter_labels[[params$sfx]] else params$sfx), collapse = ' | ')`"
author: "`r params$author`"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: sandstone
params:
  input_rds: "../../07.P.compute_cellchat/results/vars/so_w_cellchat.rds"
  author: "Giorgio Gonnella"
  pair: "A_C_vs_NC"
  sfx:  "pfilt"
  filter_mode: "pfilt"
  pval_threshold: 0.05
  output_dir: "../results"
  output_href_prefix: "../results"
  pair_definitions: !r list(A_C_vs_NC = c("A_C", "A_NC"), `6m_C_vs_NC` = c("6m_C", "6m_NC"), NC_A_vs_6m = c("A_NC", "6m_NC"), C_A_vs_6m = c("A_C", "6m_C"))
  pair_labels: !r list(A_C_vs_NC = "Acute - Chronic vs Non-Chronic", `6m_C_vs_NC` = "6 months - Chronic vs Non-Chronic", NC_A_vs_6m = "Non-Chronic - Acute vs 6 months", C_A_vs_6m = "Chronic - Acute vs 6 months")
  filter_labels: !r list(all = "unfiltered", pfilt = "p-value filtered")
  report_title_prefix: "Chikv"
  group_celltype: !r c(rep("B cells", 5), rep("CD4+ T cells", 8), rep("CD8+ T cells", 4), rep("GD T cells", 2), rep("MAIT cells", 1), rep("NK cells", 1), rep("Dendritic cells", 2), rep("Monocytes", 3), rep("Progenitor cells", 1))
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)

library(baia.cellchat)
library(Seurat)
library(CellChat)
library(patchwork)
library(ggplot2)
library(circlize)
library(ComplexHeatmap)
library(grid)
library(rmdreportdeck)
reportdeck_setup()
knitr::opts_chunk$set(cache=F)

if (is.null(params$sfx) || !nzchar(params$sfx)) params$sfx <- params$filter_mode
if (is.null(params$pair_definitions))
  stop("Pass pair_definitions to rmarkdown::render().", call. = FALSE)
if (!params$pair %in% names(params$pair_definitions))
  stop("Unknown pair: ", params$pair)
if (!params$sfx %in% c("all", "pfilt"))
  stop("Unknown sfx: ", params$sfx)

rds_key <- params$pair

conds  <- params$pair_definitions[[params$pair]]
nm1    <- conds[1]; nm2 <- conds[2]
sfx    <- params$sfx
thresh <- if (sfx == "pfilt") params$pval_threshold else 1
pfx    <- paste0(params$pair, ".", sfx)

plots_outdir <- file.path(params$output_dir, params$pair, sfx, "plots")
dir.create(plots_outdir, recursive = TRUE, showWarnings = FALSE)

fig_counter <- baia.cellchat::cellchat_fig_counter()
```

# Goal

Summarize CellChat cell–cell communication results for a given condition pair and filter setting.
Displays overview heatmaps, interaction count bar charts, signaling role heatmaps, chord diagrams,
rank-net plots, and bubble plots comparing signaling between conditions, with statistical tests
on per-sample values where applicable.

```{r load_so}
so <- readRDS(file = params$input_rds)
cc                  <- so@misc$cellchat_filtered[[sfx]]
cc_all              <- so@misc$cellchat_filtered$all
merged_cc           <- so@misc$cellchat_merged[[sfx]][[rds_key]]
grouped             <- so@misc$cellchat_grouped[[sfx]]
grouped_all         <- so@misc$cellchat_grouped$all
grouped_merged      <- so@misc$cellchat_grouped_merged[[sfx]][[rds_key]]
cellchat_per_sample <- so@misc$cellchat_per_sample
```

# Single condition overviews

## Overview heatmaps

```{r overview, results='asis'}
baia.cellchat::cellchat_results_overview_section(
  cc = cc, cc_all = cc_all, nm1 = nm1, nm2 = nm2,
  sfx = sfx, pfx = pfx, plots_outdir = plots_outdir,
  fig_counter = fig_counter
)
```

# Comparison: `r nm1` vs `r nm2`

## Pathway ranking (rankNet)

```{r ranknet, fig.width=12, fig.height=7, results='asis'}
baia.cellchat::cellchat_results_ranknet_section(
  cc = cc, cc_all = cc_all, merged_cc = merged_cc,
  nm1 = nm1, nm2 = nm2, sfx = sfx, thresh = thresh,
  pfx = pfx, plots_outdir = plots_outdir, params = params,
  fig_counter = fig_counter
)
```

## Signaling role heatmaps

```{r signaling_role_heatmaps, fig.width=12, fig.height=9, results='asis'}
baia.cellchat::cellchat_results_signaling_role_section(
  cc = cc, cc_all = cc_all, nm1 = nm1, nm2 = nm2,
  sfx = sfx, pfx = pfx, plots_outdir = plots_outdir, params = params,
  fig_counter = fig_counter
)
```

## Comparison plots

```{r comparison_plots, fig.width=12, fig.height=10, results='asis'}
baia.cellchat::cellchat_results_comparison_section(
  cc = cc, cc_all = cc_all, merged_cc = merged_cc,
  cellchat_per_sample = cellchat_per_sample,
  nm1 = nm1, nm2 = nm2, sfx = sfx, thresh = thresh,
  pfx = pfx, plots_outdir = plots_outdir, params = params,
  fig_counter = fig_counter
)
```

# Grouped cell types

## Grouped overview heatmaps

```{r grouped_overview, results='asis'}
baia.cellchat::cellchat_results_grouped_overview_section(
  grouped = grouped, grouped_all = grouped_all,
  nm1 = nm1, nm2 = nm2, sfx = sfx,
  pfx = pfx, plots_outdir = plots_outdir,
  fig_counter = fig_counter
)
```

## Grouped comparison

```{r grouped_comparison_plots, fig.width=12, fig.height=10, results='asis'}
baia.cellchat::cellchat_results_grouped_comparison_section(
  grouped = grouped, grouped_all = grouped_all, grouped_merged = grouped_merged,
  cellchat_per_sample = cellchat_per_sample, cc_all = cc_all,
  nm1 = nm1, nm2 = nm2, sfx = sfx, thresh = thresh,
  pfx = pfx, plots_outdir = plots_outdir, params = params,
  group_celltype = params$group_celltype,
  fig_counter = fig_counter
)
```

# Session info

```{r session_info, echo=FALSE}
sessionInfo()
```
