---
title: "Chikv | Post-processing after integration"
author: "Giorgio Gonnella"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: sandstone
params:
  analysis: "1"
  step: "03.P.integration"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=F, warning=F, echo=F)
withr::with_dir(here::here(), { source("renv/activate.R") })

library(Seurat)
library(RColorBrewer)

analysis_root <- here::here(paste0("analysis-", params$analysis,"/"))

this_step_dir <- paste0(analysis_root, "steps/", params$step, "/")
if (!dir.exists(this_step_dir)) {
  stop("Current step directory does not exist: ", this_step_dir)
}

var_dir <- paste0(this_step_dir, "results/vars/")
if (!dir.exists(var_dir)) {
  stop("Variable directory does not exist: ", var_dir)
}
```

```{r load, echo=F}
so <- readRDS(file = paste0(var_dir, "integrated_so.raw.rds"))
so$clinical_condition = ifelse(so$Chronicity == "True", "Chronic", "Non-Chronic")
```

# Select clustering resolution

Select the clustering resolution to be used for the downstream analysis.

```{r selected_resolution, echo=TRUE}
so@misc$sel_res <- "1"
so@misc$sel_res_col <- paste0("harmsnn_res.", so@misc$sel_res)
```

# Filters

## Remove uncharacterized cells

```{r remove_unchar}
so$unchar_ct <-so$pruned.labels_singleR_fine == "unchar"
so <- subset(so, unchar_ct == FALSE)
so$unchar_ct <- NULL
```

## Remove small cell types

Remove cells classified in cell types with less than 10 cells

```{r remove_celltype_outliers, echo=T}
min_cells_per_celltype <- 10
celltype_sizes <- table(so$pruned.labels_singleR_fine)
small_celltypes <- names(celltype_sizes[celltype_sizes < min_cells_per_celltype])
so$in_small_celltype <-so$pruned.labels_singleR_fine %in% small_celltypes
so <- subset(so, in_small_celltype == FALSE)
```

## Remove small clusters

Remove cells clustered in clusters with less than 10 cells

```{r remove_small_clusters, echo=T}
min_cluster_size = 10
cluster_sizes <- table(so@meta.data[[so@misc$sel_res_col]])
small_clusters <-
  as.integer(names(cluster_sizes[cluster_sizes < min_cluster_size]))
so$in_small_cluster <-
  so@meta.data[[so@misc$sel_res_col]] %in% small_clusters
so <- subset(so, in_small_cluster == FALSE)
so$in_small_cluster <- NULL

so@meta.data[[so@misc$sel_res_col]] <-
  droplevels(so@meta.data[[so@misc$sel_res_col]])
```

# Cell types

## Group and sort fine celltype labels

Create groups of fine celltype labels (CD4, CD8 and others) and sort them in
a meaningful order

```{r celltype_grps, echo=TRUE}
so@misc$celltype_grp <- list()
so@misc$celltype_grp[["B cells"]] <-
  c("Naive B cells", "Switched memory B cells",
    "Non-switched memory B cells", "Exhausted B cells", "Plasmablasts")
so@misc$celltype_grp[["CD4+ T cells"]] <-
  c("Naive CD4 T cells", "Th1 cells", "Th1/Th17 cells", "Th17 cells",
    "Th2 cells", "Follicular helper T cells", "T regulatory cells",
    "Terminal effector CD4 T cells")
so@misc$celltype_grp[["CD8+ T cells"]] <-
  c("Naive CD8 T cells", "Central memory CD8 T cells",
    "Terminal effector CD8 T cells", "Effector memory CD8 T cells")
so@misc$celltype_grp[["CD4-/CD8- T cells"]] <-
  c("Gamma-delta T cells", "Vd2 gd T cells",
    "Non-Vd2 gd T cells", "MAIT cells")
so@misc$celltype_grp[["NK cells"]] <-
  c("Natural killer cells")
so@misc$celltype_grp[["Dendritic cells"]] <-
  c("Plasmacytoid dendritic cells", "Myeloid dendritic cells")
so@misc$celltype_grp[["Monocytes"]] <-
  c("Intermediate monocytes", "Classical monocytes", "Non classical monocytes")
so@misc$celltype_grp[["Granulocytes"]] <-
  c("Low-density basophils")
so@misc$celltype_grp[["Progenitor cells"]] <-
  c("Progenitor cells")

so$celltype <-
  factor(so$pruned.labels_singleR_fine,
         levels = c(so@misc$celltype_grp[["B cells"]],
                    so@misc$celltype_grp[["CD4+ T cells"]],
                    so@misc$celltype_grp[["CD8+ T cells"]],
                    so@misc$celltype_grp[["CD4-/CD8- T cells"]],
                    so@misc$celltype_grp[["NK cells"]],
                    so@misc$celltype_grp[["Dendritic cells"]],
                    so@misc$celltype_grp[["Monocytes"]],
                    so@misc$celltype_grp[["Granulocytes"]],
                    so@misc$celltype_grp[["Progenitor cells"]]
                    ))

so$celltype <- droplevels(so$celltype)

so$cell_group <- NA
for (group in names(so@misc$celltype_grp)) {
  so$cell_group <-
    ifelse(so$celltype %in%
             so@misc$celltype_grp[[group]],
           group, so$cell_group)
}

so@misc$cell_lineage <- list()
so@misc$cell_lineage[["Lymphoid"]] <-
  c(so@misc$celltype_grp[["B cells"]],
    so@misc$celltype_grp[["CD4+ T cells"]],
    so@misc$celltype_grp[["CD8+ T cells"]],
    so@misc$celltype_grp[["CD4-/CD8- T cells"]],
    so@misc$celltype_grp[["NK cells"]])
so@misc$cell_lineage[["Myeloid"]] <-
  c(so@misc$celltype_grp[["Dendritic cells"]],
    so@misc$celltype_grp[["Monocytes"]],
    so@misc$celltype_grp[["Granulocytes"]])
so@misc$cell_lineage[["Other"]] <-
   so@misc$celltype_grp[["Granulocytes"]]

so$cell_lineage <- NA
for (group in names(so@misc$cell_lineage)) {
  so$cell_lineage <-
    ifelse(so$celltype %in% so@misc$cell_lineage[[group]],
           group, so$cell_lineage)
}
```

# Abbreviated fine celltype labels

```{r abbreviate_cell_types, echo=TRUE}
so@misc$celltype.abbrev <- levels(so$celltype)
names(so@misc$celltype.abbrev) <- levels(so$celltype)

so@misc$celltype.abbrev[[
    "Naive CD4 T cells"]] <- "CD4+ Naive"
so@misc$celltype.abbrev[[
    "Naive CD8 T cells"]] <- "CD8+ Naive"
so@misc$celltype.abbrev[[
    "Naive B cells"]] <- "Naive B"
so@misc$celltype.abbrev[[
    "Natural killer cells"]] <- "NK cells"
so@misc$celltype.abbrev[[
    "Classical monocytes"]] <- "Class.Monoc."
so@misc$celltype.abbrev[[
    "T regulatory cells"]] <- "Tregs"
so@misc$celltype.abbrev[[
    "Th1/Th17 cells"]] <- "Th1/Th17"
so@misc$celltype.abbrev[[
    "Vd2 gd T cells"]] <- "Vd2 GD"
so@misc$celltype.abbrev[[
    "Gamma-delta T cells"]] <- "GD"
so@misc$celltype.abbrev[[
    "Th2 cells"]] <- "Th2"
so@misc$celltype.abbrev[[
    "Non-Vd2 gd T cells"]] <- "Non-Vd2 GD"
so@misc$celltype.abbrev[[
    "Terminal effector CD8 T cells"]] <- "CD8+ TEMRA"
so@misc$celltype.abbrev[[
    "Central memory CD8 T cells"]] <- "CD8+ CM"
so@misc$celltype.abbrev[[
    "Th17 cells"]] <- "Th17"
so@misc$celltype.abbrev[[
    "Follicular helper T cells"]] <- "Tfh"
so@misc$celltype.abbrev[[
    "MAIT cells"]] <- "MAIT"
so@misc$celltype.abbrev[[
    "Th1 cells"]] <- "Th1"
so@misc$celltype.abbrev[[
    "Intermediate monocytes"]] <- "Interm.Monoc."
so@misc$celltype.abbrev[[
    "Switched memory B cells"]] <- "Sw.mem.B"
so@misc$celltype.abbrev[[
    "Effector memory CD8 T cells"]] <- "CD8+ EM"
so@misc$celltype.abbrev[[
    "Non-switched memory B cells"]] <- "Non-sw.mem.B"
so@misc$celltype.abbrev[[
    "Myeloid dendritic cells"]] <- "Myel.dendr."
so@misc$celltype.abbrev[[
    "Exhausted B cells"]] <- "Exh.B"
so@misc$celltype.abbrev[[
    "Terminal effector CD4 T cells"]] <- "CD4+ TEMRA"
so@misc$celltype.abbrev[[
    "Non classical monocytes"]] <- "Non-class.Monoc."
so@misc$celltype.abbrev[[
    "Plasmacytoid dendritic cells"]] <- "Plasmac.dendr."
so@misc$celltype.abbrev[[
    "Progenitor cells"]] <- "Progenitor"
so@misc$celltype.abbrev[[
    "Plasmablastas"]] <- "Plasmabl."
so@misc$celltype.abbrev[[
    "Low-density basophils"]] <- "Ld Basoph."

so$celltype.abbrev <-
  factor(as.vector(so@misc$celltype.abbrev[so$celltype]),
    levels = so@misc$celltype.abbrev[levels(so$celltype)])
```

# Cell types combined with TimePoint

Create new columns ``celltype.tp`` and ``celltype.abbrev.tp`` that combine the
cell type and the TimePoint. The order shall be the same of celltype and
celltype.abbrev with the two TimePoints just after each other ("ct-A", "ct-6m",
for each celltype "ct"). Also create a column ``cellgroup.tp`` that combines the
cell group and the TimePoint, with the same order ("cg-A", "cg-6m", for each
cellgroup "cg"). Also create the ``so@misc$celltype_grp.tp`` list containing
the cell groups for each of the celltype.tp, similar to `so@misc$celltype_grp`.

```{r}
so$celltype.tp <- paste0(so$celltype, "-", so$TimePoint)
so$celltype.tp <- factor(so$celltype.tp,
  levels = paste0(so@misc$celltype, "-", so$TimePoint),
  labels = paste0(so@misc$celltype, "-", so$TimePoint))

so$celltype.abbrev.tp <- paste0(so$celltype.abbrev, "-", so$TimePoint)
so$celltype.abbrev.tp <- factor(so$celltype.abbrev.tp,
  levels = paste0(so@misc$celltype.abbrev, "-", so$TimePoint),
  labels = paste0(so@misc$celltype.abbrev, "-", so$TimePoint))

so$cell_group.tp <- paste0(so$cell_group, "-", so$TimePoint)
so$cell_group.tp <- factor(so$cell_group.tp,
  levels = paste0(so@misc$cell_group, "-", so$TimePoint),
  labels = paste0(so@misc$cell_group, "-", so$TimePoint))

so@misc$celltype_grp.tp <- list()
for (group in names(so@misc$celltype_grp)) {
  so@misc$celltype_grp.tp[[group]] <-
    paste0(so@misc$celltype_grp[[group]], "-", so$TimePoint)
}
```

# Color schemes

Define color schemes for the main cell type labels, fine cell type labels,
clinical conditions, samples, clusters and cell cycle phases, in order to have
consistent colors across all plots.

The colors are saved in some columns of the `so@misc` slot:
- `mctype_cols` for the main cell type labels
- `fctype_cols` for the fine cell type labels
- `cond_cols` for the clinical conditions
- `sample_cols` for the samples
- `cluster_cols` for the clusters
- `phase_cols` for the cell cycle phases
- `celltype_groups_bkg` for plots background divided by celltype group

```{r color_scheme_for_main_labels}
so@misc$mctype_cols <-
  colorRampPalette(brewer.pal(6, "Accent"))(
    length(unique(so$pruned.labels_singleR_main)))
names(so@misc$mctype_cols) <-
  levels(so$pruned.labels_singleR_main)
```

```{r color_scheme_for_fine_labels}
so@misc$colors$celltypes <- c(
  # B cells (shades of blue)
  "Naive B cells" = "#1f78b4",           # blue
  "Switched memory B cells" = "#6baed6", # light blue
  "Non-switched memory B cells" = "#0c61ff", # dark blue/cyan
  "Exhausted B cells" = "#084594",       # very dark blue
  "Plasmablasts" = "#9ecae1",            # pale blue

  # CD4+ T cells (shades of green)
  "Naive CD4 T cells" = "#238b45",       # green
  "Th1 cells" = "#006d2c",               # dark green
  "Th1/Th17 cells" = "#74c476",          # light green
  "Th17 cells" = "#41ab5d",              # medium green
  "Th2 cells" = "#a1d99b",               # pale green
  "Follicular helper T cells" = "#31a354", # another shade of green
  "T regulatory cells" = "#c7ffc0",      # very pale green
  "Terminal effector CD4 T cells" = "#005a32", # very dark green

  # CD8+ T cells (shades of red)
  "Naive CD8 T cells" = "#e31a1c",       # red
  "Central memory CD8 T cells" = "#fc9272", # light red
  "Effector memory CD8 T cells" = "#fb6a4a",  # medium red
  "Terminal effector CD8 T cells" = "#66181d", # dark red

  # CD4-/CD8- T cells (shades of purple)
  "Gamma-delta T cells" = "#6a51a3",    # purple
  "Vd2 gd T cells" = "#6050aa",         # purple
  "Non-Vd2 gd T cells" = "#9e9ac8",     # light purple
  "MAIT cells" = "#4a3183",             # dark purple

  # Natural killer cells (pink)
  "Natural killer cells" = "#dd3497",    # pink

  # Monocytes (shades of brown)
  "Classical monocytes" = "#8c510a",     # brown
  "Intermediate monocytes" = "#4f410d",  # light brown
  "Non classical monocytes" = "#543005", # dark brown

  # Dendritic cells (shades of orange)
  "Plasmacytoid dendritic cells" = "#fe9929", # orange
  "Myeloid dendritic cells" = "#ec7014",      # dark orange

  # Granulocytes (yellow)
  "Low-density basophils" = "#ffff33",    # bright yellow

  # Progenitor cells (gray)
  "Progenitor cells" = "#2f2f2f"         # dark gray
)
```

```{r color_scheme_for_conditions}
so@misc$cond_cols <- c(
  "Non-Chronic" = "darkblue",
  "Chronic" = "darkred"
)
```

```{r color_scheme_for_samples}
so@misc$sample_cols <-
  colorRampPalette(brewer.pal(8, name = "Accent"))(
                  length(unique(so$sample)))
names(so@misc$sample_cols) <- levels(so$sample)
```

```{r color_scheme_for_clusters}
g.b = so@misc$sel_res_col
so@misc$cluster_cols <-
  colorRampPalette(brewer.pal(8, name = "Set1"))(
              length(unique(so@meta.data[[g.b]])))
names(so@misc$cluster_cols) <-
  levels(so@meta.data[[g.b]])
```

```{r color_scheme_for_phases}
so@misc$phase_cols <- c("#00AFBB", "#E7B800", "#FC4E07")
names(so@misc$phase_cols) <- c("G1", "G2M", "S")
```

```{r color_scheme_for_plots_background_divided_by_celltype_group}
so@misc$colors$celltype_groups_bkg <- c(
  "B cells" = "#cce5ff",              # Light blue
  "CD4+ T cells" = "#ddffdd",         # Light green
  "CD8+ T cells" = "#ffcc99",         # Light orange
  "CD4-/CD8- T cells" = "#e6ccff",    # Light purple
  "NK cells" = "#ffe6e6",             # Light pink
  "Dendritic cells" = "#ffffcc",      # Light yellow
  "Monocytes" = "#eeff99",            # Light green/yellow
  "Granulocytes" = "#aabbbb",         # Light gray
  "Progenitor cells" = "#fdd3d3",     # Light red
  "Uncharacterized" = "#f0f0f0"       # Light gray
)
```

## Saving data

```{r save, echo=TRUE}
saveRDS(so, paste0(var_dir, "integrated_so.rds"))
```

## Session info

```{r, echo=F}
sessionInfo()
```
