---
title: "Chikv | Merging and Integration"
author: "Giorgio Gonnella"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: sandstone
params:
  analysis: "1"
  step: "03.P.integration"
  input_step: "02.P.filter"
  input_sfx: "_so_filtered.rds"
  samples_sheet: "metadata/samples.tsv"
  sample_id_col: "SampleID"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message=F, warning=F, echo=F)
withr::with_dir(here::here(), { source("renv/activate.R") })

library(harmony)
library(Seurat)

analysis_root <- here::here(paste0("analysis-", params$analysis,"/"))

input_step_dir <- paste0(analysis_root, "steps/", params$input_step, "/")
input_var_dir <- paste0(input_step_dir, "results/vars/")

if (!dir.exists(input_var_dir)) {
  stop("Input variable directory does not exist: ", input_var_dir)
}

this_step_dir <- paste0(analysis_root, "steps/", params$step, "/")
if (!dir.exists(this_step_dir)) {
  stop("Current step directory does not exist: ", this_step_dir)
}

var_dir <- paste0(this_step_dir, "results/vars/")
dir.create(var_dir, showWarnings = FALSE, recursive = TRUE)

samples_sheet <- here::here(params$samples_sheet)

helpers_dir <- here::here("helpers/")
source(paste0(helpers_dir, "merging_helpers.R"))
source(paste0(helpers_dir, "clustering_helpers.R"))
source(paste0(helpers_dir, "samples_sheet.R"))
```

# Load Seurat Objects

```{r read input data}
samples_metadata <- read_samples_sheet(samples_sheet,
                                       params$sample_id_col)
sampleIDs <- get_sample_IDs(samples_metadata,
                            column=params$sample_id_col)

so_list = list()
for (sampleID in sampleIDs) {
  filename = paste0(input_var_dir, sampleID, params$input_sfx)
  so_list[[sampleID]] <- readRDS(filename)
  so_list[[sampleID]]$BatchID <- samples_metadata$BatchID[samples_metadata[[params$sample_id_col]] == sampleID]
  so_list[[sampleID]]$Chronicity <- samples_metadata$Chronicity[samples_metadata[[params$sample_id_col]] == sampleID]
  so_list[[sampleID]]$ByCRFs <- samples_metadata$ByCRFs[samples_metadata[[params$sample_id_col]] == sampleID]
  so_list[[sampleID]]$Age <- samples_metadata$Age[samples_metadata[[params$sample_id_col]] == sampleID]
  so_list[[sampleID]]$Sex <- samples_metadata$Sex[samples_metadata[[params$sample_id_col]] == sampleID]
  so_list[[sampleID]]$SequencingYear <- samples_metadata$SequencingYear[samples_metadata[[params$sample_id_col]] == sampleID]
  so_list[[sampleID]]$sample <- sampleID
  so_list[[sampleID]]$PatientID <- samples_metadata$PatientID[samples_metadata[[params$sample_id_col]] == sampleID]
  so_list[[sampleID]]$TimePoint <- samples_metadata$TimePoint[samples_metadata[[params$sample_id_col]] == sampleID]
  so_list[[sampleID]]$FlowcellID <- samples_metadata$FlowcellID[samples_metadata[[params$sample_id_col]] == sampleID]
  so_list[[sampleID]]$Lane <- samples_metadata$Lane[samples_metadata[[params$sample_id_col]] == sampleID]
  so_list[[sampleID]]$RunID <- samples_metadata$RunID[samples_metadata[[params$sample_id_col]] == sampleID]
}
```

# Merge objects into a single Seurat object

```{r merge}
so <- merge(x = so_list[[1]], y = so_list[-1], project = "chikv")
```

## Normalization

```{r normalization, echo=TRUE}
so <- NormalizeData(object = so, assay = "RNA")
```

## Selection of highly variable genes

```{r hvg_selection, echo=TRUE}
so <- FindVariableFeatures(object = so, assay = "RNA",
                           nfeatures = 2500, selection.method = "vst")
```

Removing unwanted source of variation in the highly variable genes

```{r correcting_hvg, echo=TRUE}
vf <- VariableFeatures(so)

unwanted_tcr <- grep(pattern = "^TR[AB][VC]", vf, value = TRUE)
unwanted_ig <- c(grep(pattern = "^IG[HLK][VDJ]", vf, value = TRUE),
              c("IGHM", "IGHD", "IGHE", "IGHA[12]", "IGHG[1234]",
                "IGKC", "IGLC[1234567]", "AC233755.1"))
unwanted_mt <- grep(pattern = "^MT", vf, value = TRUE)
unwanted_ribo <- grep(pattern = "^RP[LS]", vf, value = TRUE)
unwanted_vf <- c(unwanted_tcr, unwanted_ig, unwanted_mt, unwanted_ribo)
hvg_corr <- vf[!vf %in% unwanted_vf]
```

## Scaling data

```{r scaling, echo=TRUE}
so <- ScaleData(object = so,
            features = vf,
            vars.to.regress = c("nCount_RNA", "percent.mito"),
            assay = "RNA")
```

## Linear dimension reduction (PCA)

```{r pca, echo=TRUE}
so <- RunPCA(object = so, features = hvg_corr,
             npcs = 50, assay = "RNA")
```

## Non-linear dim reduction (UMAP)

```{r umap_pre, echo=TRUE}
myDims <- 1:40
so <- RunUMAP(object = so, dims = myDims,
              reduction = "pca", reduction.name = "umap")
```

## Clustering

```{r clustering_pre, echo=TRUE}
reS <- c(1.0)
so <- FindNeighbors(object = so, dims = myDims)
so <- FindClusters(object = so, resolution = reS)
so <- factor_resolution(so)
```

# Harmony integration

In the integration step, the Harmony algorithm is used to integrate the data
from the different samples. It is a method which aims to correct for batch
effects while preserving the biological signal.

```{r run_harmony, echo = TRUE}

set.seed(42)
so <- RunHarmony(object = so,
                group.by.vars=c("orig.ident"),
                reduction = "pca", reduction.save="harmony")
```

## UMAP

```{r setting_umap_parameters, echo=TRUE}
myDims <- 1:40
```

```{r umap , message=FALSE, warning=FALSE}
set.seed(42)
so <- RunUMAP(object = so, dims = myDims,
              reduction = "harmony", reduction.name = "umapharm")
```

## Clustering

```{r setting_clustering_parameters, echo = TRUE}
reS <- c(0.1, 0.5, 1.0, 2.5, 5.0)
```

```{r clustering, message = FALSE, warning = FALSE, echo = TRUE}
set.seed(42)
so <- FindNeighbors(object = so, dims = myDims,
                    reduction = "harmony", graph.name = "harmsnn")
so <- FindClusters(object = so, resolution = reS,
                   graph.name = "harmsnn")
so <- factor_resolution(so)
```

# Saving data

```{r save, echo = TRUE}
saveRDS(so, file = paste0(var_dir, "integrated_so.raw.rds"))
```

# Session info

```{r session_info}
sessionInfo()
```
