---
title: "Chikv | Results after Harmony integration"
author: "Giorgio Gonnella"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: sandstone
params:
  analysis: "1"
  input_step: "03.P.integration"
  step: "03.V.integrated"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=F, warning=F, echo=F, cache=F)
withr::with_dir(here::here(), { source("renv/activate.R") })

library(Seurat)
library(viridis)
library(ggplot2)
library(RColorBrewer)
library(rmdreportdeck)
reportdeck_setup()
knitr::opts_chunk$set(cache=F)

analysis_root <- here::here(paste0("analysis-", params$analysis,"/"))

input_step_dir <- paste0(analysis_root, "steps/", params$input_step, "/")
input_var_dir <- paste0(input_step_dir, "results/vars/")
if (!dir.exists(input_var_dir)) {
  stop("Input variable directory does not exist: ", input_var_dir)
}

this_step_dir <- paste0(analysis_root, "steps/", params$step, "/")
if (!dir.exists(this_step_dir)) {
  stop("Current step directory does not exist: ", this_step_dir)
}

plots_outdir <- paste0(this_step_dir, "results/plots/")
dir.create(plots_outdir, recursive = T, showWarnings = F)
tables_outdir <- paste0(this_step_dir, "results/tables/")
dir.create(tables_outdir, recursive = T, showWarnings = F)

helpers_dir <- here::here("helpers/")
source(paste0(helpers_dir, "combined_proportion_plot.R"))
```

```{r load, echo = FALSE}
so <- readRDS(paste0(input_var_dir, "integrated_so.rds"))

umap_df <- as.data.frame(Embeddings(so, reduction = "umapharm"))
colnames(umap_df)[1:2] <- c("UMAP_1", "UMAP_2")
umap_df <- cbind(
  data.frame(cell_id = rownames(umap_df), check.names = FALSE),
  umap_df[, 1:2, drop = FALSE],
  so@meta.data[, c("sample", "Phase", so@misc$sel_res_col,
                   "pruned.labels_singleR_fine"), drop = FALSE]
)
colnames(umap_df)[colnames(umap_df) == so@misc$sel_res_col] <- "cluster"
```

# Goal

Summarize the integrated Harmony object after dimensional reduction and annotation.
Display UMAP by sample, phase, cluster, and cell type;
quantify sample, cluster, phase, and cell-type proportions.

# UMAP

## Overall UMAP plot (Fig V03.01)

```{r umap, fig.align = "center", echo = FALSE, fig.width = 12, fig.height = 6, results = "asis"}
plot_obj <- DimPlot(object = so, reduction = "umapharm",
                    group.by = "sample",
                    cols = so@misc$sample_cols) +
  ggtitle("UMAP plot, colored by sample")
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03.01",
  suffix = "umap_by_sample",
  width = 12,
  height = 6,
  data = umap_df,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle, tsv_label = "Download TSV")), "\n")
print(plot_obj)
```

## UMAP plots by sample (Fig V03.02)

```{r umap_by_sample, fig.align = "center", echo = FALSE, fig.width = 12, fig.height = 12, results = "asis"}
plot_obj <- DimPlot(object = so, reduction = "umapharm",
                    group.by = "sample", cols = so@misc$sample_cols,
                    split.by = "sample", ncol = 3) +
  NoLegend() +
  ggtitle("UMAP plot, splitted by sample")
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03.02",
  suffix = "umap_split_by_sample",
  width = 12,
  height = 12,
  data = umap_df,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle, tsv_label = "Download TSV")), "\n")
print(plot_obj)
```

## UMAP plot by phase (Fig V03.03)

Basic metrics projected onto the UMAP space

```{r umap_basic_metrics, fig.align = "center", echo = FALSE, results = "asis"}
plot_obj <- DimPlot(object = so, reduction = "umapharm", group.by = "Phase",
                    cols = so@misc$phase_cols) +
  ggtitle("UMAP plot, colored by phase")
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03.03",
  suffix = "umap_by_phase",
  width = 7,
  height = 5,
  data = umap_df,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle, tsv_label = "Download TSV")), "\n")
print(plot_obj)
```

# Clustering

The selected clustering resolution is `r so@misc$sel_res`.

## UMAP by cluster (Fig V03.04)

```{r umap_clustering, fig.align = "center", fig.width = 12, fig.height = 6, echo = FALSE, results = "asis"}
plot_obj <- DimPlot(object = so, reduction = "umapharm",
                    group.by = so@misc$sel_res_col, label = TRUE,
                    cols = so@misc$cluster_cols) +
  ggtitle("UMAP plot, colored by cluster")
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03.04",
  suffix = "umap_by_cluster",
  width = 12,
  height = 6,
  data = umap_df,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle, tsv_label = "Download TSV")), "\n")
print(plot_obj)
```

## Cluster composition by sample (Fig V03.05)

```{r clusters_by_sample, fig.align = "center", fig.width = 12, fig.height = 6, echo = FALSE, results = "asis"}
sample_vs_cl <- as.data.frame(
  round(prop.table(table(so@meta.data[[so@misc$sel_res_col]],
                         so$sample), margin = 1)*100,
        digits = 2)
)
colnames(sample_vs_cl)[1:2] <- c("Cluster", "SampleID")
levels(sample_vs_cl$Cluster) <-
  paste("Cl", levels(sample_vs_cl$Cluster), sep = "_")
plot_obj <- ggplot(sample_vs_cl, aes(x = Cluster, y = Freq, fill = SampleID))+
  geom_bar(stat = "identity", alpha = .98)+
  coord_flip()+
  theme_minimal()+
  scale_fill_manual(values = so@misc$sample_cols)
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03.05",
  suffix = "cluster_composition_by_sample",
  width = 12,
  height = 6,
  data = sample_vs_cl,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle, tsv_label = "Download TSV")), "\n")
print(plot_obj)
```

## Sample composition by cluster (Fig V03.06)

```{r samples_by_cluster, fig.align = "center", fig.width = 12, fig.height = 6, echo = FALSE, results = "asis"}
cols <- so@misc$cluster_cols
names(cols) <- paste0("Cl_", names(so@misc$cluster_cols))
cl_vs_sample <- as.data.frame(
  round(prop.table(table(so$sample,
                         so@meta.data[[so@misc$sel_res_col]]), margin = 1)*100,
        digits = 2)
)
colnames(cl_vs_sample)[1:2] <- c("SampleID", "Cluster")
levels(cl_vs_sample$Cluster) <-
  paste("Cl", levels(cl_vs_sample$Cluster), sep = "_")
plot_obj <- ggplot(cl_vs_sample, aes(x = SampleID, y = Freq, fill = Cluster))+
  geom_bar(stat = "identity", alpha = .98)+
  coord_flip()+
  theme_minimal()+
  scale_fill_manual(values = cols)
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03.06",
  suffix = "sample_composition_by_cluster",
  width = 12,
  height = 6,
  data = cl_vs_sample,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle, tsv_label = "Download TSV")), "\n")
print(plot_obj)
```

# Cell types

## UMAP by cell type (Fig V03.07)

```{r umap_singleR_fine, fig.width = 12, fig.height = 6, echo = FALSE, results = "asis"}
plot_obj <- DimPlot(object = so, reduction = "umapharm",
                    group.by = "pruned.labels_singleR_fine",
                    cols = so@misc$colors$celltypes, label = F, repel = TRUE) +
  ggtitle("UMAP plot, colored by cell type")
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03.07",
  suffix = "umap_by_singleR_fine_celltype",
  width = 12,
  height = 6,
  data = umap_df,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle, tsv_label = "Download TSV")), "\n")
print(plot_obj)
```

## UMAP by cell type, split by sample (Fig V03.08)

```{r umap_singleR_by_sample_fine, fig.width = 12, fig.height = 12, echo = FALSE, results = "asis"}
plot_obj <- DimPlot(object = so, reduction = "umapharm",
                    group.by = "pruned.labels_singleR_fine",
                    cols = so@misc$colors$celltypes, label = F, repel = TRUE,
                    split.by = "sample", ncol = 3) +
  NoLegend() +
  ggtitle("UMAP plot, colored by cell type, splitted by sample")
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03.08",
  suffix = "umap_by_singleR_fine_celltype_split_by_sample",
  width = 12,
  height = 12,
  data = umap_df,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle, tsv_label = "Download TSV")), "\n")
print(plot_obj)
```

## Cell type proportions per sample (Tab V03.09)

```{r table_fine_celltypes_per_sample, echo = FALSE, results = "asis"}
tbl <- as.data.frame.matrix(round(prop.table(table(so$sample,
  so$pruned.labels_singleR_fine), margin = 2)*100, digits = 2))
tbl <- data.frame(sample_id = rownames(tbl), tbl, check.names = FALSE)
cat(as.character(report_asset_bar(
  report_table_name("V03.09"),
  report_tsv_download_link(
    tbl,
    report_table_name("V03.09", "fine_celltypes_per_sample.tsv"),
    "Download TSV"
  )
)), "\n")
knitr::kable(tbl)
```

## Cell type proportions per sample (Fig V03.10)

```{r plot_fine_celltypes_per_sample, fig.width = 12, fig.height = 6, echo = FALSE, results = "asis"}
celltype_df_f <- as.data.frame(round(prop.table(table(so$sample,
  so$pruned.labels_singleR_fine), margin = 2)*100, digits = 2))
colnames(celltype_df_f) <- c("sample_id", "cell_type", "percentage")
plot_obj <- ggplot(celltype_df_f, aes(x = cell_type, y = percentage))+
  geom_bar(stat = "identity", aes(fill = sample_id))+
  scale_fill_manual(values = so@misc$sample_cols)+
  theme_bw()+
  coord_flip()+
  scale_x_discrete(limits = rev(levels(celltype_df_f$cell_type)))
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03.10",
  suffix = "fine_celltypes_per_sample",
  width = 12,
  height = 6,
  data = celltype_df_f,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle, tsv_label = "Download TSV")), "\n")
print(plot_obj)
```

## Sample proportions per cell type (Tab V03.11)

```{r table_samples_per_fine_celltype, echo = FALSE, results = "asis"}
tbl <- as.data.frame.matrix(round(prop.table(
    table(so$pruned.labels_singleR_fine,
          so$sample), margin = 2)*100, digits = 2))
tbl <- data.frame(cell_type = rownames(tbl), tbl, check.names = FALSE)
cat(as.character(report_asset_bar(
  report_table_name("V03.11"),
  report_tsv_download_link(
    tbl,
    report_table_name("V03.11", "samples_per_fine_celltype.tsv"),
    "Download TSV"
  )
)), "\n")
knitr::kable(tbl)
```

## Sample proportions per cell type (Fig V03.12)

```{r plot_samples_per_fine_celltypes, fig.width = 12, echo = FALSE, results = "asis"}
celltype_df_fr <- as.data.frame(round(prop.table(
      table(so$pruned.labels_singleR_fine,
            so$sample), margin = 2)*100, digits = 2))
colnames(celltype_df_fr) <- c("cell_type", "sample_id", "percentage")
plot_obj <- ggplot(celltype_df_fr, aes(x = sample_id, y = percentage))+
  geom_bar(stat = "identity", aes(fill = cell_type))+
  scale_fill_manual(values = so@misc$colors$celltypes)+
  theme_bw()+
  coord_flip()+
  scale_x_discrete(limits = rev(levels(celltype_df_fr$sample_id)))
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03.12",
  suffix = "samples_per_fine_celltype",
  width = 12,
  height = 6,
  data = celltype_df_fr,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle, tsv_label = "Download TSV")), "\n")
print(plot_obj)
```

# Cell types by cluster

## SingleR assignments by cluster (Tab V03.13)

```{r singleR_vs_clustering_tables, results = "asis", echo = FALSE}
singleR_cluster_counts <- as.data.frame.matrix(
  table(so@meta.data[[so@misc$sel_res_col]],
        so$pruned.labels_singleR_fine)
)
singleR_cluster_counts <- data.frame(
  cluster = rownames(singleR_cluster_counts),
  singleR_cluster_counts,
  check.names = FALSE
)
singleR_cluster_row_pct <- as.data.frame.matrix(addmargins(round(prop.table(
  table(so@meta.data[[so@misc$sel_res_col]],
        so$pruned.labels_singleR_fine),
  margin = 1)*100, digits = 2), margin = 2))
singleR_cluster_row_pct <- data.frame(
  cluster = rownames(singleR_cluster_row_pct),
  singleR_cluster_row_pct,
  check.names = FALSE
)
singleR_cluster_col_pct <- as.data.frame.matrix(addmargins(round(prop.table(
  table(so$pruned.labels_singleR_fine,
        so@meta.data[[so@misc$sel_res_col]]),
  margin = 1)*100, digits = 2), margin = 2))
singleR_cluster_col_pct <- data.frame(
  cell_type = rownames(singleR_cluster_col_pct),
  singleR_cluster_col_pct,
  check.names = FALSE
)
cat(as.character(report_asset_bar(
  report_table_name("V03.13"),
  report_tsv_download_link(singleR_cluster_counts,
                           report_table_name("V03.13", "singleR_vs_cluster_counts.tsv"),
                           "Counts TSV"),
  report_tsv_download_link(singleR_cluster_row_pct,
                           report_table_name("V03.13", "singleR_vs_cluster_cluster_pct.tsv"),
                           "Cluster % TSV"),
  report_tsv_download_link(singleR_cluster_col_pct,
                           report_table_name("V03.13", "singleR_vs_cluster_celltype_pct.tsv"),
                           "Cell type % TSV")
)), "\n")
print(knitr::kable(singleR_cluster_counts))
print(knitr::kable(singleR_cluster_row_pct))
print(knitr::kable(singleR_cluster_col_pct))
```

## Cell type composition by cluster (Fig V03.14)

```{r singleR_vs_clustering_plots1, fig.width = 12, fig.height = 6, echo = FALSE, results = "asis"}
df <- as.data.frame(round(prop.table(table(so@meta.data[[so@misc$sel_res_col]],
        so$pruned.labels_singleR_fine), margin = 2)*100,
        digits = 2))
colnames(df) <- c("cluster", "cell_type", "percentage")
plot_obj <- ggplot(df, aes(x = cell_type, y = percentage))+
  geom_bar(stat = "identity", aes(fill = cluster))+
  theme_bw()+
  coord_flip()+
  scale_x_discrete(limits = rev(levels(df$cell_type))) +
  scale_fill_manual(values = so@misc$cluster_cols)
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03.14",
  suffix = "singleR_celltype_composition_by_cluster",
  width = 12,
  height = 6,
  data = df,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle, tsv_label = "Download TSV")), "\n")
print(plot_obj)
```

## Cluster composition by cell type (Fig V03.15)

```{r singleR_vs_clustering_plots2, fig.width=12, fig.height=6, echo=F, results = "asis"}
df2 <- as.data.frame(round(prop.table(table(
         so$pruned.labels_singleR_fine,
         so@meta.data[[so@misc$sel_res_col]]), margin = 2)*100, digits = 2))
colnames(df2) <- c("cell_type", "cluster", "percentage")
plot_obj <- ggplot(df2, aes(x = cluster, y = percentage))+
  geom_bar(stat = "identity", aes(fill = cell_type))+
  theme_bw()+
  coord_flip()+
  scale_x_discrete(limits = rev(levels(df2$cluster))) +
  scale_fill_manual(values = so@misc$colors$celltypes)
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03.15",
  suffix = "cluster_composition_by_singleR_celltype",
  width = 12,
  height = 6,
  data = df2,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle, tsv_label = "Download TSV")), "\n")
print(plot_obj)
```

# Phase by cluster

## Phase vs. cluster counts and proportions (Tab V03.16)

```{r phase_vs_clustering_tables, results = "asis", echo = FALSE}
phase_cluster_counts <- as.data.frame.matrix(
  table(so@meta.data[[so@misc$sel_res_col]],
        so$Phase)
)
phase_cluster_counts <- data.frame(
  cluster = rownames(phase_cluster_counts),
  phase_cluster_counts,
  check.names = FALSE
)
phase_cluster_row_pct <- as.data.frame.matrix(addmargins(round(prop.table(
  table(so@meta.data[[so@misc$sel_res_col]],
        so$Phase),
  margin = 1)*100, digits = 2), margin = 2))
phase_cluster_row_pct <- data.frame(
  cluster = rownames(phase_cluster_row_pct),
  phase_cluster_row_pct,
  check.names = FALSE
)
phase_cluster_col_pct <- as.data.frame.matrix(addmargins(round(prop.table(
  table(so$Phase,
        so@meta.data[[so@misc$sel_res_col]]),
  margin = 1)*100, digits = 2), margin = 2))
phase_cluster_col_pct <- data.frame(
  phase = rownames(phase_cluster_col_pct),
  phase_cluster_col_pct,
  check.names = FALSE
)
cat(as.character(report_asset_bar(
  report_table_name("V03.16"),
  report_tsv_download_link(phase_cluster_counts,
                           report_table_name("V03.16", "phase_vs_cluster_counts.tsv"),
                           "Counts TSV"),
  report_tsv_download_link(phase_cluster_row_pct,
                           report_table_name("V03.16", "phase_vs_cluster_cluster_pct.tsv"),
                           "Cluster % TSV"),
  report_tsv_download_link(phase_cluster_col_pct,
                           report_table_name("V03.16", "phase_vs_cluster_phase_pct.tsv"),
                           "Phase % TSV")
)), "\n")
print(knitr::kable(phase_cluster_counts))
print(knitr::kable(phase_cluster_row_pct))
print(knitr::kable(phase_cluster_col_pct))
```

## Phase composition by cluster (Fig V03.17)

```{r phase_vs_clustering_plots1, fig.width=12, fig.height=6, echo=F, results = "asis"}
df <- as.data.frame(round(prop.table(table(so@meta.data[[so@misc$sel_res_col]],
        so$Phase), margin = 2)*100,
        digits = 2))
colnames(df) <- c("cluster", "phase", "percentage")
plot_obj <- ggplot(df, aes(x = phase, y = percentage))+
  geom_bar(stat = "identity", aes(fill = cluster))+
  theme_bw()+
  coord_flip()+
  scale_x_discrete(limits = rev(levels(df$phase))) +
  scale_fill_manual(values = so@misc$cluster_cols)
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03.17",
  suffix = "phase_composition_by_cluster",
  width = 12,
  height = 6,
  data = df,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle, tsv_label = "Download TSV")), "\n")
print(plot_obj)
```

## Cluster composition by phase (Fig V03.18)

```{r phase_vs_clustering_plots2, fig.width=12, fig.height=6, echo=F, results = "asis"}
df2 <- as.data.frame(round(prop.table(table(
         so$Phase,
         so@meta.data[[so@misc$sel_res_col]]), margin = 2)*100, digits = 2))
colnames(df2) <- c("phase", "cluster", "percentage")
plot_obj <- ggplot(df2, aes(x = cluster, y = percentage))+
  geom_bar(stat = "identity", aes(fill = phase))+
  theme_bw()+
  coord_flip()+
  scale_x_discrete(limits = rev(levels(df2$cluster))) +
  scale_fill_manual(values = so@misc$phase_cols)
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03.18",
  suffix = "cluster_composition_by_phase",
  width = 12,
  height = 6,
  data = df2,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle, tsv_label = "Download TSV")), "\n")
print(plot_obj)
```

# Phase by cell type

## Phase vs. cell type counts and proportions (Tab V03.19)

```{r phase_vs_fine_celltype_tables, results="asis", echo=F}
phase_celltype_counts <- as.data.frame.matrix(
  table(so$pruned.labels_singleR_fine,
        so$Phase)
)
phase_celltype_counts <- data.frame(
  cell_type = rownames(phase_celltype_counts),
  phase_celltype_counts,
  check.names = FALSE
)
phase_celltype_row_pct <- as.data.frame.matrix(addmargins(round(prop.table(
  table(so$pruned.labels_singleR_fine,
        so$Phase),
  margin = 1)*100, digits = 2), margin = 2))
phase_celltype_row_pct <- data.frame(
  cell_type = rownames(phase_celltype_row_pct),
  phase_celltype_row_pct,
  check.names = FALSE
)
phase_celltype_col_pct <- as.data.frame.matrix(addmargins(round(prop.table(
  table(so$Phase,
        so$pruned.labels_singleR_fine),
  margin = 1)*100, digits = 2), margin = 2))
phase_celltype_col_pct <- data.frame(
  phase = rownames(phase_celltype_col_pct),
  phase_celltype_col_pct,
  check.names = FALSE
)
cat(as.character(report_asset_bar(
  report_table_name("V03.19"),
  report_tsv_download_link(phase_celltype_counts,
                           report_table_name("V03.19", "phase_vs_fine_celltype_counts.tsv"),
                           "Counts TSV"),
  report_tsv_download_link(phase_celltype_row_pct,
                           report_table_name("V03.19", "phase_vs_fine_celltype_celltype_pct.tsv"),
                           "Cell type % TSV"),
  report_tsv_download_link(phase_celltype_col_pct,
                           report_table_name("V03.19", "phase_vs_fine_celltype_phase_pct.tsv"),
                           "Phase % TSV")
)), "\n")
print(knitr::kable(phase_celltype_counts))
print(knitr::kable(phase_celltype_row_pct))
print(knitr::kable(phase_celltype_col_pct))
```

## Phase composition by cell type (Fig V03.20)

```{r phase_vs_fine_celltype_plots1, fig.width=12, fig.height=6, echo=F, results = "asis"}
df <- as.data.frame(round(prop.table(table(
        so$pruned.labels_singleR_fine,
        so$Phase), margin = 2)*100,
        digits = 2))
colnames(df) <- c("cell_type", "phase", "percentage")
plot_obj <- ggplot(df, aes(x = phase, y = percentage))+
  geom_bar(stat = "identity", aes(fill = cell_type))+
  theme_bw()+
  coord_flip()+
  scale_x_discrete(limits = rev(levels(df$phase))) +
  scale_fill_manual(values = so@misc$colors$celltypes)
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03.20",
  suffix = "phase_composition_by_fine_celltype",
  width = 12,
  height = 6,
  data = df,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle, tsv_label = "Download TSV")), "\n")
print(plot_obj)
```

## Cell type composition by phase (Fig V03.21)

```{r phase_vs_fine_celltype_plots2, fig.width=12, fig.height=6, echo=F, results = "asis"}
df2 <- as.data.frame(round(prop.table(table(
         so$Phase,
         so$pruned.labels_singleR_fine), margin = 2)*100, digits = 2))
colnames(df2) <- c("phase", "cell_type", "percentage")
plot_obj <- ggplot(df2, aes(x = cell_type, y = percentage))+
  geom_bar(stat = "identity", aes(fill = phase))+
  theme_bw()+
  coord_flip()+
  scale_x_discrete(limits = rev(levels(df2$cell_type))) +
  scale_fill_manual(values = so@misc$phase_cols)
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03.21",
  suffix = "fine_celltype_composition_by_phase",
  width = 12,
  height = 6,
  data = df2,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle, tsv_label = "Download TSV")), "\n")
print(plot_obj)
```

# Session info

```{r, echo=F}
sessionInfo()
```
