---
title: "Chikv | Results after Harmony integration | UMAP highlighting single cell types"
author: "Giorgio Gonnella"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: sandstone
params:
  analysis: "1"
  input_step: "03.P.integration"
  step: "03.V.integrated"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=F, warning=F, echo=F, cache=F)
withr::with_dir(here::here(), { source("renv/activate.R") })

library(Seurat)
library(viridis)
library(ggplot2)
library(RColorBrewer)
library(rmdreportdeck)
reportdeck_setup()
knitr::opts_chunk$set(cache=F)

analysis_root <- here::here(paste0("analysis-", params$analysis,"/"))

input_step_dir <- paste0(analysis_root, "steps/", params$input_step, "/")
input_var_dir <- paste0(input_step_dir, "results/vars/")
if (!dir.exists(input_var_dir)) {
  stop("Input variable directory does not exist: ", input_var_dir)
}

this_step_dir <- paste0(analysis_root, "steps/", params$step, "/")
if (!dir.exists(this_step_dir)) {
  stop("Current step directory does not exist: ", this_step_dir)
}

plots_outdir <- paste0(this_step_dir, "results/plots/")
dir.create(plots_outdir, recursive = T, showWarnings = F)
tables_outdir <- paste0(this_step_dir, "results/tables/")
dir.create(tables_outdir, recursive = T, showWarnings = F)

```

```{r load, echo = FALSE}
so <- readRDS(paste0(input_var_dir, "integrated_so.rds"))

umap_df <- as.data.frame(Embeddings(so, reduction = "umapharm"))
colnames(umap_df)[1:2] <- c("UMAP_1", "UMAP_2")
umap_df <- cbind(
  data.frame(cell_id = rownames(umap_df), check.names = FALSE),
  umap_df[, 1:2, drop = FALSE],
  so@meta.data[, c("sample", "Phase", so@misc$sel_res_col,
                   "pruned.labels_singleR_fine"), drop = FALSE]
)
colnames(umap_df)[colnames(umap_df) == so@misc$sel_res_col] <- "cluster"
```

# Goal

Visualize the Harmony-integrated UMAP embedding with one SingleR fine cell type highlighted per panel.
Each panel displays all cells in grey with the cells of a given type overlaid in dark blue,
allowing visual confirmation of each cell type's spatial distribution on the UMAP.

# UMAP plot colored by cell type

```{r umap_by_celltype, fig.width = 12, fig.height = 6, echo = FALSE, results = "asis"}
plot_obj <- DimPlot(object = so, reduction = "umapharm",
                    group.by = "pruned.labels_singleR_fine",
                    cols = so@misc$colors$celltypes, label = FALSE, repel = TRUE) +
  ggtitle("UMAP plot, colored by cell type")
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V03S.00",
  suffix = "umap_by_celltype",
  width = 12,
  height = 6,
  plot_dir = plots_outdir,
  data_dir = tables_outdir
)
cat(as.character(report_bundle_asset_bar(bundle$stem, bundle)), "\n")
print(plot_obj)
```

# UMAP highlighting single cell types

```{r single_cell_types_umap, fig.width = 12, fig.height = 6, echo = FALSE, results = "asis"}
cell_types <- sort(unique(so$pruned.labels_singleR_fine))
items <- vector("list", length(cell_types))

for (i in seq_along(cell_types)) {
  ct <- cell_types[[i]]
  highl <- WhichCells(so, expression = (pruned.labels_singleR_fine == ct))
  plt <- DimPlot(so, reduction = "umapharm",
        cells.highlight = highl, cols.highlight = "darkblue",
        cols = "grey") +
    ggtitle(ct) + theme(legend.position = "none")
  ct_slug <- tolower(gsub("(^-+|-+$)", "", gsub("[^A-Za-z0-9]+", "-", ct)))
  bundle <- report_plot_bundle(
    plt,
    block_prefix = paste0("V03S.", sprintf("%02d", i)),
    suffix = ct_slug,
    width = 12,
    height = 6,
    plot_dir = plots_outdir,
    data_dir = tables_outdir
  )
  items[[i]] <- report_item_panel(
    title = paste0(ct, " (", length(highl), " cells) (Fig V03S.", sprintf("%02d", i), ")"),
    asset_bar = report_bundle_asset_bar(bundle$stem, bundle),
    report_item_plot(plt, width = 12, height = 6)
  )
}

cat(as.character(report_loop_section(items)))
```

# Session info

```{r, echo=F}
sessionInfo()
```
